%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Task 4 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\task{Transistor-based AC/DC converter (single-phase AFE)}

\taskGerman{Transistorbasierter AC/DC-Umrichter}

An electric-drive test bench is connected to a single-phase grid via a full-bridge transistor AFE rectifier.
The converter should operate with approximately unity power factor and a stable DC-link voltage.
During testing, increased DC-link ripple is observed and the PWM strategy is questioned.

\begin{germanblock}
    Ein elektrischer Antriebsprüfstand ist über einen einphasigen Vollbrücken-Transistor-Gleichrichter
    an das Netz angeschlossen. Der Umrichter soll mit näherungsweise Leistungsfaktor Eins und stabiler
    DC-Zwischenkreisspannung betrieben werden. Während der Tests wird eine erhöhte Zwischenkreiswelligkeit
    beobachtet und die PWM-Strategie wird hinterfragt.
\end{germanblock}

\input{fig/Task04/Single-phase_DC_Inverter.tex}
\begin{table}[ht]
    \centering
    \begin{tabular}{lll}
        \toprule
        Parameter             & Symbol            & Value                     \\
        \midrule
        Grid voltage (RMS)    & $U_\mathrm{g}$             & $\SI{230}{\volt}$         \\
        Grid frequency        & $f$               & $\SI{50}{\hertz}$         \\
        Grid-side inductance  & $L$               & $\SI{2.5}{\milli\henry}$  \\
        DC-link voltage       & $U_\mathrm{dc}$          & $\SI{450}{\volt}$         \\
        Active power setpoint & $P$               & $\SI{3.0}{\kilo\watt}$    \\
        DC-link capacitor     & $C_\mathrm{dc}$ & $\SI{2.2}{\milli\farad}$  \\
        Switching period      & $T_\mathrm{s}$             & $\SI{100}{\micro\second}$ \\
        PWM reference signal  & $s^*(t)$          & sinusoidal                \\
        \bottomrule
    \end{tabular}
    \caption{Nominal operating parameters of the single-phase AFE rectifier.}
    \label{tab:task04_parameters}
\end{table}

\subtask{For approximately unity power factor operation, determine the required grid current RMS value and
    the corresponding grid current peak value.}{2}
\subtaskGerman{Bestimmen Sie für den Betrieb mit näherungsweise Leistungsfaktor Eins den erforderlichen
    Effektivwert des Netzstroms sowie den zugehörigen Spitzenwert des Netzstroms.}
\begin{solutionblock}
    Grid current at unity power factor is:
    $$P=U_\mathrm{g} I_\mathrm{g} \Rightarrow I_\mathrm{g}=\frac{\SI{3000}{\watt}}{\SI{230}{\volt}}=\SI{13.04}{\ampere}.$$
    $$\hat{i}_\mathrm{g}=\sqrt{2}I_\mathrm{g}=\SI{18.44}{\ampere}. $$
\end{solutionblock}
\subtask{Calculate the required converter input-voltage amplitude $\hat{u}_2$ and verify
    whether the nominal DC-link voltage is feasible.}{1}
\subtaskGerman{Berechnen Sie aus der komplexen Zeigerbeziehung die erforderliche Amplitude der Umrichtereingangsspannung $\hat{u}_2$
    und prüfen Sie, ob die nominelle DC-Zwischenkreisspannung ausreicht.}
\begin{solutionblock}
    Required converter input-voltage amplitude yields:
    $$\hat{u}_2=\sqrt{\hat{u}_\mathrm{g}^2+(\omega L\hat{i}_\mathrm{g})^2}
        =\sqrt{(\SI{325.27}{\volt})^2+\bigl(\SI{314.16}{\per\second}\cdot\SI{2.5}{\milli\henry}\cdot\SI{18.44}{\ampere}\bigr)^2}
        =\SI{325.59}{\volt}. $$
    With $U_{\mathrm{dc}}=\SI{450}{\volt}>\hat{u}_2$, the boost rectifier condition is fulfilled.
\end{solutionblock}
\subtask{Determine the DC-link voltage oscillation amplitude $\hat{u}_\mathrm{C}$.}{4}
\subtaskGerman{Bestimmen Sie die Amplitude der DC-Zwischenkreis-Spannungsoszillation $\hat{u}_\mathrm{C}$.}
\begin{solutionblock}
    Calculating the DC-link ripple requires the apparent power at the converter input:
    $$Q_2=\omega L I_\mathrm{g}^2=\SI{314.16}{\per\second}\cdot\SI{2.5}{\milli\henry}\cdot(\SI{13.04}{\ampere})^2=\SI{133.7}{\volt\ampere}.$$
    $$S_2=\sqrt{P^2+Q_2^2}=\sqrt{(\SI{3000}{\watt})^2+(\SI{133.7}{\volt\ampere})^2}=\SI{3003}{\volt\ampere}. $$
    From lecture Eq. (6.23), the DC-link current ripple amplitude is approximately
    $$\hat{i}_\mathrm{C}\approx\frac{S_2}{U_\mathrm{dc}}=\frac{3003}{450}=\SI{6.67}{\ampere}. $$
    Hence, the DC-link voltage ripple amplitude results in
    $$\hat{u}_\mathrm{C}\approx\frac{\hat{i}_\mathrm{C}}{2\omega C_\mathrm{dc}}=\frac{\SI{6.67}{\ampere}}{2\cdot\SI{314.16}{\per\second}\cdot\SI{2.2}{\milli\farad}}=\SI{4.82}{\volt}.$$
    Alternatively,
    The average DC-link current follows from the active power balance:
    $$I_{0}=\frac{P}{U_\mathrm{dc}}=\frac{\SI{3000}{\watt}}{\SI{450}{\volt}}=\SI{6.67}{\ampere}. $$
    The DC-link voltage ripple amplitude is
    $$\hat{u}_\mathrm{C} = \hat{u}_1 \approx \frac{\hat{i}_\mathrm{C}}{2\omega C_\mathrm{dc}} = \frac{I_0}{2\omega C_\mathrm{dc}}\sqrt{1 + \left(\frac{\omega L U_\mathrm{dc}}{U^2_\mathrm{g}}\right)^2}.$$
    Substituting the values yields
    $$\hat{u}_\mathrm{C} = \frac{\SI{6.67}{\ampere}}{2\cdot\SI{314.16}{\per\second}\cdot\SI{2.2}{\milli\farad}}\sqrt{1 + \left(\frac{\SI{314.16}{\per\second}\cdot\SI{2.5}{\milli\henry}\cdot\SI{450}{\volt}}{\SI{230}{\volt}^2}\right)^2}=\SI{4.82}{\volt}.$$

\end{solutionblock}
\subtask{The engineering team suspects that the choice of PWM strategy contributes to the observed DC-link ripple.
    Using a sinusoidal reference $s^*(t)=\hat{s}\sin(\omega t)$ and a triangular carrier $c(t)\in[-1,1]$, sketch the switching states of both complementary PWM and interleaved PWM over one fundamental electrical period ($\omega t\in[0,2\pi]$) in the provided templates.
    Finally, explain which strategy produces a lower AC-side current ripple in $L$ and why this reduces stress on the DC-link capacitor $C_{\mathrm{dc}}$.}{4}
\subtaskGerman{Das Engineering-Team vermutet, dass die Wahl der PWM-Strategie zur beobachteten DC-Zwischenkreiswelligkeit beiträgt.
    Verwenden Sie ein sinusförmiges Referenzsignal $s^*(t)=\hat{s}\sin(\omega t)$ und einen Dreiecksträger $c(t)\in[-1,1]$, und skizzieren Sie die Schaltzustände beider Strategien (Komplementär-PWM und Überlagerungs-PWM) über eine elektrische Grundperiode ($\omega t\in[0,2\pi]$) in den bereitgestellten Vorlagen.
    Erläutern Sie abschließend, welche Strategie eine geringere netzstromseitige Stromwelligkeit in $L$ erzeugt und warum dies die Belastung des DC-Zwischen\-kreis\-kon\-den\-sa\-tors $C_{\mathrm{dc}}$ verringert.}
\input{fig/Task04/Fig_subtask4.5_pwm_template_complementary.tex}
\input{fig/Task04/Fig_subtask4.5_pwm_template_interleaved.tex}
\begin{solutionblock}
    Given:
    The modulation reference is sinusoidal, $s^*(t)=\hat{s}\sin(\omega t)$. The sketches are shown over one fundamental electrical period ($\omega t\in[0,2\pi]$).

    1) Complementary PWM
    $$s_1(t)=\begin{cases}
            +1, & s^*(t)>c(t)    \\
            -1, & s^*(t)\le c(t)
        \end{cases},\qquad s_2(t)=-s_1(t).$$
    The equivalent converter state is
    $$s_{\mathrm{cs}}(t)=\frac{s_1(t)-s_2(t)}{2}=s_1(t)\in\{+1,-1\}. $$

    2) Interleaved PWM
    $$s_1(t)=\begin{cases}
            +1, & s^*(t)>c(t)    \\
            -1, & s^*(t)\le c(t)
        \end{cases},\qquad
        s_2(t)=\begin{cases}
            +1, & -s^*(t)>c(t)    \\
            -1, & -s^*(t)\le c(t)
        \end{cases}. $$
    The equivalent three-level state is
    $$s_{\mathrm{is}}(t)=\frac{s_1(t)-s_2(t)}{2}\in\{+1,0,-1\}. $$
    A correct sketch over one electrical period shows transition instants at the intersections of $c(t)$ with $s^*(t)$ and $-s^*(t)$.

    What must be visible in the plots:
    \begin{itemize}
        \item Complementary PWM: one carrier $c(t)$, one reference $s^*(t)$, states $s_1,s_2$, and resulting $s\in\{+1,-1\}$.
        \item Interleaved PWM: one carrier $c(t)$, both references $s^*(t)$ and $-s^*(t)$, states $s_1,s_2$, and resulting $s\in\{+1,0,-1\}$.
    \end{itemize}

    \input{fig/Task04/sFig_subtask4.5_pwm_solution_complementary.tex}
    \input{fig/Task04/sFig_subtask4.5_pwm_solution_interleaved.tex}

    Comparison and impact on DC-link ripple:

    Complementary PWM produces a two-level converter output state ($\pm1$), while interleaved PWM produces a three-level output state ($+1,0,-1$).
    The additional zero level in interleaved PWM reduces the effective voltage steps across $L$ and therefore reduces the AC-side current ripple.
    Since the high-frequency capacitor current is caused by the switching ripple power of the converter,
    the smaller inductor current ripple of interleaved PWM reduces the current stress of $C_{\mathrm{dc}}$
    and therefore lowers the DC-link voltage ripple $\Delta U_{dc,pp}$ observed in the previous subtask.
    Interleaved PWM is therefore the preferable strategy.
    \vfill
\end{solutionblock}
