%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Examtask 2: SEPIC converter as smart voltage stabilizer for bicycle dynamo %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\task{SEPIC as smart voltage stabilizer for bicycle dynamo}

\taskGerman{SEPIC als intelligente Spannungsstabilisierung für Fahrraddynamos}

A SEPIC is to be designed for voltage stabilization in a bicycle dynamo.
A control circuit regulates the converter's duty cycle to ensure a stable output voltage.
The dynamo provides a voltage of $\SI{3}{\volt} \ldots \SI{10}{\volt}$. The output voltage is to be stabilized at $\SI{5.1}{\volt}$.

\begin{germanblock}
    Für ein Fahraddynamo soll ein SEPIC als Spannungsstabilisierung dimensioniert werden.
    Eine Reglungsschaltung regelt das Tastverhältnis des Konverter für eine stabile Ausgangsspannung.
    Der Dynamo liefert eine Spannung von $\SI{3}{\volt} \ldots \SI{10}{\volt}$. Die Ausgangsspannung soll auf $\SI{5.1}{\volt}$ stabilisiert werden.
\end{germanblock}

\input{fig/Task02/figSEPICConverter}

\begin{table}[ht]
    \centering  % Zentriert die Tabelle
    \begin{tabular}{llll}
        \toprule
        \multicolumn{2}{l}{\textbf{General parameters:}}                                                                                                                          \\
        Input voltage:       & $U_{\mathrm{1}} = \SI{3}{\volt} \ldots \SI{10}{\volt}$ & Diode forward voltage: & $U_{\mathrm{D},\mathrm{f}} = \SI{0.8}{\volt}$                    \\
        Output voltage:      & $U_{\mathrm{2}} = \SI{5}{\volt}$                       & Output current range:  & $I_{\mathrm{2}} = \SI{50}{\milli\ampere} \ldots \SI{1}{\ampere}$ \\
        Switching frequency: & $f_\mathrm{s} = \SI{100}{\kilo\hertz}$                 &                        &                                                                  \\
        \midrule
        \multicolumn{4}{l}{\quad The diode forward voltage must be taken into account.}                                                                                           \\
        \multicolumn{4}{l}{\quad The voltage of capacitor $C_{\mathrm{1}}$ can be considered as constant for the specific duty cycles.}                                           \\
        \multicolumn{4}{l}{\quad The output voltage can be considered as constant except of subtask 2.4.}                                                                         \\
        \bottomrule
    \end{tabular}
    \caption{Parameters of the circuit.}  % Beschriftung der Tabelle
    \label{table:ex02_Parameters of the circuit}
\end{table}


\subtask{What are the minimum and maximum duty cycles that the control circuit has to command to maintain the desired output voltage given the varying input voltage?}{0}
\subtaskGerman{Welche minimalen und maximalen Tastverhältnisse muss die Regelungsschaltung vorgeben, um die gewünschte Ausgangsspannung bei der variierenden Eingangsspannung zu halten?}

% Solution of subtask
\begin{solutionblock}
    The duty cycle  for the SEPIC is calculated by
    \begin{equation*}
        \frac{U_\mathrm{out}}{U_\mathrm{in}}=\frac{D}{1-D}.
    \end{equation*}
    Solving this equation with respect to $D$ leads to
    \begin{equation*}
        D=\frac{U_\mathrm{out}}{U_\mathrm{in}+U_\mathrm{out}}
    \end{equation*}
    For the circuit the voltages are like follow:
    \begin{align*}
        U_\mathrm{out} & =U_\mathrm{2}+U_\mathrm{D} \\
        U_\mathrm{in}  & =U_\mathrm{1}
    \end{align*}
    Takein this in accournt the duty cycles yields
    For the circuit the voltages are like follow:
    \begin{align*}
        D              & =\frac{U_\mathrm{2}+U_\mathrm{D}}{U_\mathrm{in}+U_\mathrm{2}+U_\mathrm{D}}                 \\
        D_\mathrm{min} & =\frac{\SI{5}{\volt}+\SI{0.8}{\volt}}{\SI{3}{\volt}+\SI{5}{\volt}+\SI{0.8}{\volt}}= 0.37   \\
        D_\mathrm{max} & =\frac{\SI{5}{\volt}+\SI{0.8}{\volt}}{\SI{10}{\volt}+\SI{5}{\volt}+\SI{0.8}{\volt}}= 0.66.
    \end{align*}
\end{solutionblock}

\subtask{Which operation point is decisive for the calculation of minimal $L_{\mathrm{2}}$,
    to avoid discontinuous conduction mode (DCM) within the specified operating range? Calculate the minimal required inductances $L_2$ for this case.}{0}
\subtaskGerman{Welcher Arbeitspunkt ist bei der Berechnung der minimalen Induktivität $L{\mathrm{2}}$ ausschlaggebend,
    um den Lückebetrieb innerhalb des definierten Arbeitsbereichs zu vermeiden? Berechnen Sie die minimal erforderliche Induktivität $L_2$ für diesen Fall.}

% Solution of subtask
\begin{solutionblock}
    The inductance $L_\mathrm{2}$ provide the energy to the load while transistor is blocking.
    The maximum $\Delta i_\mathrm{L2}$ occurs at the maximum transistor off-time, which corresponds to the minimum duty cycle.
    At this duty cycle, the converter first enters BCM mode. It means for all larger duty cycles the converter are in CCM mode.
    In general BCM mode occurs when the minimum average inductor current equals the half of the inductor current ripple within the considered time range.
    The critical operation point for $L_\mathrm{2}$ is at minimum duty cycle $D_\mathrm{min}$ and minimum load current at $I_\mathrm{2,min} = \SI{50}{\milli\ampere}$.
    If the transistor is on, no current is supplied to the output. Therefore, the correspondent average current flowing during $T_\mathrm{off, max}$ must be taken into account:
    \begin{equation*}
        \overline{i}_\mathrm{L2,off} = \frac{I_\mathrm{2,min}}{1-D_\mathrm{min}}.
    \end{equation*}
    Using this current $\Delta i_\mathrm{L_2}$ is calculated by
    \begin{equation*}
        \Delta i_\mathrm{L_2}= 2 \cdot \overline{i}_\mathrm{L2,off} = 2 \cdot \frac{I_\mathrm{2,min}}{1-D_\mathrm{min}}
    \end{equation*}
    The ripple current $\Delta i_\mathrm{L_2}$ yields
    \begin{equation*}
        \Delta i_\mathrm{L2}=\frac{U_\mathrm{L2,off} \cdot T_\mathrm{off, max}}{L2}
        =\frac{(U_\mathrm{2} + U_\mathrm{D}) \cdot (1-D_\mathrm{min})}{L_\mathrm{2} \cdot f_\mathrm{s}}.
    \end{equation*}
    This leads to
    \begin{equation*}
        2 \cdot \frac{I_\mathrm{2,min}}{1-D_\mathrm{min}}=\frac{(U_\mathrm{2} + U_\mathrm{D}) \cdot (1-D_\mathrm{min})}{L_\mathrm{2} \cdot f_\mathrm{s}}.
    \end{equation*}
    Solving this equation with respect to $L$ yields
    \begin{equation*}
        L_\mathrm{2} = \frac{(U_\mathrm{2} + U_\mathrm{D}) \cdot (1-D_\mathrm{min})^2}{2 \cdot \Delta i_\mathrm{L2} \cdot f_\mathrm{s}}
        = \frac{(\SI{5}{\volt} + \SI{0.8}{\volt}) \cdot (1-0.37 )^2} {2 \cdot \SI{0.1}{\ampere} \cdot \SI{100}{\kilo\hertz}}
        = \SI{230}{\micro\henry}.
    \end{equation*}
\end{solutionblock}


\subtask{To further avoid operation in DCM mode within the defined operating range, the minimum inductance  $L_{\mathrm{1}}$ must be calculated.
    For this purpose, the duty cycle at which the maximum ripple current occurs should be considered.
    Calculate the minimum required inductance $L_{\mathrm{1}}$ for this case.}{0}
\subtaskGerman{Um weiterhin den Lückbetrieb innerhalb des definierten Arbeitsbereichs zu vermeiden, soll die minimalen Induktivität $L_{\mathrm{1}}$ berechnet werden.  Dazu ist das Tastverhältnis zu betrachten, bei dem der maximale Ripplestrom auftritt.  Berechnen Sie die minimal erforderliche Induktivität $L_{\mathrm{1}}$ für diesen Fall.}

% Solution of subtask
\begin{solutionblock}
    The maximal ripple current occurs at $D=0.5$.
    The input voltage $U_\mathrm{1}$ is calculated by
    \begin{equation*}
        \frac{U_\mathrm{out}}{U_\mathrm{in}}=\frac{D}{1-D}.
    \end{equation*}
    Solving this equation with respect to $U_\mathrm{in}$ yields
    \begin{equation*}
        U_\mathrm{in}=\frac{(1-D) \cdot U_\mathrm{out}}{D} = \frac{(1 - 0.5) \cdot (\SI{5}{\volt}+\SI{0.8}{\volt})}{0.5}= \SI{5.8}{\volt}
    \end{equation*}
    If the transistor is on, no current is supplied to the output. Therefore, the correspondent average current flowing during $T_\mathrm{on}$ must be taken into account.
    \begin{equation*}
        \overline{i}_\mathrm{L1,on} = \frac{I_\mathrm{2,min}}{D}
    \end{equation*}
    Using this current $\Delta i_\mathrm{L1}$ is calculated by
    \begin{equation*}
        \Delta i_\mathrm{L1}= 2 \cdot \overline{i}_\mathrm{L1,on} = 2 \cdot \frac{I_\mathrm{2,min}}{D}
    \end{equation*}
    The ripple current $\Delta i_\mathrm{L2}$ yields
    \begin{equation*}
        \Delta i_\mathrm{L1}=\frac{U_\mathrm{in} \cdot T_\mathrm{on}}{L1}
        =\frac{U_\mathrm{in} \cdot D}{L_\mathrm{1} \cdot f_\mathrm{s}}.
    \end{equation*}
    This leads to
    \begin{equation*}
        2 \cdot \frac{I_\mathrm{2,min}}{D}=\frac{U_\mathrm{in} \cdot D}{L_\mathrm{1} \cdot f_\mathrm{s}}.
    \end{equation*}
    Solving this equation with respect to $L$ yields
    \begin{equation*}
        L_\mathrm{1} = \frac{U_\mathrm{in} \cdot D^2} {2 \cdot \Delta i_\mathrm{L2} \cdot f_\mathrm{s}}
        = \frac{\SI{5.8}{\volt} \cdot 0.5^2} {2 \cdot \SI{0.1}{\ampere} \cdot \SI{100}{\kilo\hertz}}
        = \SI{145}{\micro\henry}.
    \end{equation*}
\end{solutionblock}

\subtask{How large must the filter capacitor be so that, in the worst case, the voltage fluctuation is $\pm \SI{5}{\percent}$ ?}{0}
\subtaskGerman{Wie groß muss der Filterkondensator ausgelegt werden, damit im ungünstigsten Fall die Spannungsschwankung $\pm \SI{5}{\percent}$ ?}

% Solution of subtask
\begin{solutionblock}
    The defined maximum voltage drop is calculated by
    \begin{equation*}
        \Delta u_\mathrm{2} = U_\mathrm{2} \cdot 2 \cdot \SI{5}{\percent} = \SI{5}{\volt} \cdot 2  \cdot \SI{5}{\percent}= \SI{0.5}{\volt}.
    \end{equation*}
    The voltage drop occurs during the time when no current is supplied to the output. This happens while the transistor is off.
    The capacitor will be discharged with the constant current $I_\mathrm{2}$ while the transistor conducts.
    The voltage drop is calculated by
    \begin{equation*}
        \Delta u_\mathrm{2} =  \frac{I_\mathrm{2} \cdot D}{ C \cdot f_\mathrm{s}}
    \end{equation*}
    The longer the transistor conducts and the higher the output current, the greater the voltage drop.
    This corresponds to the maximum duty cycle and maximum output current.
    Solving this equation with respect to $C$ yields
    \begin{equation*}
        C = \frac{I_\mathrm{2,max} \cdot D_\mathrm{2,max}}{f_\mathrm{s} \cdot  \Delta u_\mathrm{2,max}}
        = \frac{\SI{1}{\ampere} \cdot 0.66} {\SI{100} {\kilo\hertz} \cdot  \SI{0.5}{\volt}} = \SI{13}{\micro\farad} .
    \end{equation*}
\end{solutionblock}


