%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Task 3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\task{Thyristor-based AC/DC converter (M2C)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\taskGerman{Thyristorbasierter AC/DC-Umrichter (M2C)}

In an electroplating process, a single-phase two-pulse midpoint thyristor converter (M2C)
with center-tapped transformer is used as a controllable DC source.
After product quality issues, the engineering team must verify whether the converter is operated
with suitable firing angles and acceptable grid impact.

\begin{germanblock}
In einem Galvanikprozess wird ein einphasiger, zweipulsiger Mittelpunkt-Thyristorumrichter (M2C)
mit mittig angezapftem Transformator als einstellbare DC-Quelle eingesetzt.
Nach Qualitätsproblemen muss das Engineering-Team prüfen, ob geeignete Zündwinkel eingestellt sind
und ob die Netzbelastung innerhalb zulässiger Grenzen liegt.
\end{germanblock}

\input{fig/Task03/M2C_with_output_filter.tex}

\begin{table}[ht]
    \centering
    \begin{tabular}{lll}
    \toprule
    Parameter & Symbol & Value \\
    \midrule
    Primary grid voltage (RMS) & $U_1$ & $\SI{230}{\volt}$ \\
    Grid frequency & $f_1$ & $\SI{50}{\hertz}$ \\
    Transformer turns ratio & $N_2/N_1$ & $1.0$ \\
    Firing angle (normal operation) & $\alpha_0$ & $\SI{70}{\degree}$ \\
    DCM conduction angle (low-load case) & $\beta$ & $\SI{140}{\degree}$ \\
    Average output current (CCM) & $I_2$ & $\SI{30}{\ampere}$ \\
    Required plating voltage setpoint & $\overline{u}_{2,\mathrm{set}}$ & $\SI{75}{\volt}$ \\
    \bottomrule
    \end{tabular}
    \caption{Task 3 parameters.}
    \label{tab:task03_parameters}
\end{table}

\subtask{In normal production (CCM), calculate the average output voltage $\overline{u}_{2,\mathrm{CCM}}$ for $\alpha_0=\SI{70}{\degree}$.}{2}
\subtaskGerman{Berechnen Sie im Normalbetrieb (CCM) die mittlere Ausgangsspannung $\overline{u}_{2,\mathrm{CCM}}$ für $\alpha_0=\SI{70}{\degree}$.}
\begin{solutionblock}
    \textbf{Given:} $U_1=\SI{230}{\volt}$, $N_2/N_1=1$, $\alpha_0=\SI{70}{\degree}$.

    \textbf{Auxiliary values:}
    $$\hat{u}_1=\sqrt{2}U_1=\SI{325.27}{\volt},\qquad \hat{u}_s=\frac{1}{2}\hat{u}_1=\SI{162.63}{\volt}.$$

    \textbf{Average output voltage in CCM, at $\alpha_0$}
    $$\overline{u}_{2,\mathrm{CCM}}=\hat{u}_s\frac{2}{\pi}\cos\alpha_0=\SI{103.54}{\volt}\cdot\cos(70^\circ)=\SI{35.41}{\volt}.$$
\end{solutionblock}
\subtask{To correctly set the firing angle in nominal CCM operation, determine $\alpha_\mathrm{set}$ for
$\overline{u}_{2,\mathrm{set}}=\SI{75}{\volt}$.}{0}
\subtaskGerman{Bestimmen Sie zur Einstellung des korrekten Zündwinkels im nominellen CCM-Betrieb den Winkel $\alpha_\mathrm{set}$ für
$\overline{u}_{2,\mathrm{set}}=\SI{75}{\volt}$.}
\begin{solutionblock}
    \textbf{Firing angle for $\overline{u}_{2,\mathrm{set}}$ in CCM}
    $$\overline{u}_{2,\mathrm{set}}=\hat{u}_s\frac{2}{\pi}\cos\alpha_{\mathrm{set}}\Rightarrow
    \cos\alpha_{\mathrm{set}}=\frac{\SI{75}{\volt}}{\SI{103.54}{\volt}}=0.7244,
    \quad \alpha_{\mathrm{set}}\approx\SI{43.6}{\degree}.$$
\end{solutionblock}
\subtask{In open-loop operation with fixed firing angle $\alpha_0$, explain why the converter can enter DCM at low load and why this complicates maintaining a constant plating voltage.}{0}
\subtaskGerman{Erläutern Sie beim offenen Betrieb mit festem Zündwinkel $\alpha_0$, warum der Umrichter bei kleiner Last in den DCM übergehen kann und warum dadurch eine konstante Galvanikspannung schwieriger einzuhalten ist.}
\begin{solutionblock}
    At low load, the demanded output current decreases and can reach zero before the next thyristor pair is triggered.
    This interrupts conduction and the converter changes from CCM to DCM.
    In DCM, the average output voltage depends not only on $\alpha$ but also on the load-dependent conduction interval ($\beta$),
    so a fixed firing angle can no longer guarantee a constant plating voltage.
\end{solutionblock}
\subtask{Assuming low-load operation, the converter enters DCM with $\beta=\SI{140}{\degree}$.
Calculate $\overline{u}_{2,\mathrm{DCM}}$ for $\alpha_0=\SI{70}{\degree}$ and compare both the CCM and DCM values to the required setpoint $\overline{u}_{2,\mathrm{set}}=\SI{75}{\volt}$.
Briefly explain why this operating behavior can lead to quality deviations and how to address it}{0}
\subtaskGerman{Bei Teillast geht der Umrichter in den DCM mit $\beta=\SI{140}{\degree}$ über.
Berechnen Sie $\overline{u}_{2,\mathrm{DCM}}$ für $\alpha_0=\SI{70}{\degree}$ und vergleichen Sie die Werte aus CCM und DCM mit dem Sollwert $\overline{u}_{2,\mathrm{set}}=\SI{75}{\volt}$.
Erläutern Sie kurz, warum dieses Betriebsverhalten zu Qualitätsabweichungen führen kann und wie man es adressieren kann.}
\begin{solutionblock}
    \textbf{Average output voltage in DCM, at $\alpha_0$, $\beta$}
    $$\overline{u}_{2,\mathrm{DCM}}=\hat{u}_s\frac{2}{\pi}\sin\!\left(\frac{\beta}{2}\right)\sin\!\left(\alpha_0+\frac{\beta}{2}\right)
    =\SI{103.54}{\volt}\cdot\sin(70^\circ)\sin(140^\circ)=\SI{62.50}{\volt}.$$
    Neither value reaches the required setpoint of $\SI{75}{\volt}$: in CCM the output is only $\SI{35.41}{\volt}$,
    far below target, and although DCM raises it to $\SI{62.50}{\volt}$, this is still below the setpoint
    and load-dependent, causing unpredictable plating conditions and inconsistent product quality.
    Either, choose higher inductance to avoid DCM at low load, or implement closed-loop control to adjust $\alpha$ and maintain the setpoint regardless of load variations.
  %  Therefore, closed-loop voltage control (adjusting $\alpha$ online) is required to maintain the setpoint under load variations.
\end{solutionblock}
\subtask{At this new operating point ($\alpha=\alpha_\mathrm{set}$, $I_2=\SI{30}{\ampere}$), calculate
$I_1^{(1)}$, $P_1$, $Q_1^{(1)}$, and $S_1^{(1)}$ for the fundamental component. If $\alpha_\mathrm{set}$ is unknown, use $\alpha_\mathrm{0}$ instead.}{0}
\subtaskGerman{Berechnen Sie für den neuen Betriebspunkt ($\alpha=\alpha_\mathrm{set}$, $I_2=\SI{30}{\ampere}$)
$I_1^{(1)}$, $P_1$, $Q_1^{(1)}$ und $S_1^{(1)}$ der Grundschwingung. Falls $\alpha_\mathrm{set}$ unbekannt ist, verwenden Sie stattdessen $\alpha_\mathrm{0}$.}
\begin{solutionblock}
    \textbf{Fundamental complex power at $\alpha_{\mathrm{set}}$}
    $$I_1^{(1)}=\frac{\sqrt{2}}{\pi}I_2=\SI{13.50}{\ampere}.$$
    $$S_1^{(1)}=I_2\hat{u}_{s0}=\SI{3106}{\volt\ampere}.$$
    $$P_1=I_2\hat{u}_{s0}\cos\alpha_{\mathrm{set}}=\SI{2250}{\watt},\qquad
    Q_1^{(1)}=I_2\hat{u}_{s0}\sin\alpha_{\mathrm{set}}=\SI{2142}{\volt\ampere}.$$
\end{solutionblock}
\subtask{Including harmonics, calculate total apparent power $S_1$ and total reactive power $Q_1$,
and briefly assess whether the harmonic share is significant for grid compliance.}{0}
\subtaskGerman{Berechnen Sie unter Berücksichtigung der Oberschwingungen die gesamte Scheinleistung $S_1$ und die gesamte Blindleistung $Q_1$
und bewerten Sie kurz, ob der Oberschwingungsanteil für die Netzkonformität relevant ist.}
\begin{solutionblock}
    \textbf{Including harmonics}
    $$S_1=\frac{\pi}{2\sqrt{2}}S_1^{(1)}=\SI{3450}{\volt\ampere},$$
    $$Q_1=\sqrt{S_1^2-P_1^2}=\SI{2615}{\volt\ampere}.$$
\end{solutionblock}