%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Examtask 1: Step-down converter with output filter %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\task{Step-down converter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\taskGerman{Tiefsetzsteller}
%%%%%%%%%%%
% Combination of Tasks 1.2 + 1.3 from our exercises (little bit of steady state analysis plus losses / efficiency calculation)
% Reduction to (at most) 5 or 6 subquestions, i.e. a subset of the original exercises limited to some core questions for the buck converter
% Use a completely different set of parameter / signal values to change the setting compared to the exercise tasks, maybe also add a little 'pre story' for which application the buck converter should be used
% However, the 'spirit' of this first exam task should be definitely entry friendly, i.e., provide the students with a good start into the PE exam enabling to gain confidence for the next tasks 
%%%%%%%%%%%

In industrial control systems, a $\SI{50}{\volt}$ DC power supply is commonly used to power various components.
Additionally, several sensors and servo motors require a stable $\SI{12}{\volt}$ DC power supply.
For this purpose an efficient step-down (buck) converter is required to provide high currents,
especially when multiple servo motors or actuators are operating simultaneously.

\begin{germanblock}
	In industriellen Steuerungssystemen wird häufig eine $\SI{50}{\volt}$ DC-Stromversorgung verwendet,
	um verschiedene Komponenten zu versorgen. Zusätzlich benötigen mehrere Sensoren und Servomotoren
	eine stabile $\SI{12}{\volt}$ DC-Spannung. Dafür wird ein effizienter Tiefsetzsteller benötigt,
	der hohe Ströme liefern kann – insbesondere, wenn mehrere Servomotoren oder Aktoren gleichzeitig
	in Betrieb sind.
\end{germanblock}

% figure figStepDownConverterOutputFilter
\input{fig/Task01/figStepDownConverterOutputFilter}

\begin{table}[ht]
	\centering  % Zentriert die Tabelle
	\begin{tabular}{llll}
		\toprule
		\multicolumn{2}{l}{\textbf{General parameters:}} & \multicolumn{2}{l}{\textbf{IGBT:}}                                                                                        \\
		Input voltage:                                   & $U_{\mathrm{1}} = \SI{50}{\volt}$      & Collector-emitter voltage: & $u_{\mathrm{on},\mathrm{CE}} = \SI{2}{\volt}$       \\
		Output voltage:                                  & $U_2 = \SI{12}{\volt}$                 & Switch-on losses:          & $E_{\text{on},\mathrm{D}} = \SI{20}{\micro\joule}$  \\
		Output current:                                  & $I_2 = \SI{70}{\ampere}$               & Switch-off losses:         & $E_{\text{off},\mathrm{D}} = \SI{40}{\micro\joule}$ \\
		Switching frequency:                             & $f_\mathrm{s} = \SI{100}{\kilo\hertz}$ &                            &                                                     \\
		\midrule
		\textbf{Inductance:}                             & $L = \SI{20}{\micro\henry}$            &                            &                                                     \\
		\multicolumn{4}{c}{The diode is considered as ideal and the filter capacitor is $C \rightarrow \infty$.}                                                                     \\
		\bottomrule
	\end{tabular}
	\caption{Parameters of the circuit.}  % Beschriftung der Tabelle
	\label{table:ex01_Parameters of the circuit}
\end{table}

\subtask{At what duty cycle $D$ should the step-down converter be operated?
	Calculate and sketch the voltage $u_\mathrm{L}(t)$ and  current $i_\mathrm{L}(t)$ over 2 periods.}{4}%
\begin{hintblock}
	The voltage drop across the transistor must be taken into account.
\end{hintblock}

\subtaskGerman{Bei welchem Tastverhältnis $D$ sollte der Tiefsetzsteller betrieben werden?
	Skizzieren Sie die Spannung $u_\mathrm{L}(t)$ und den Strom $i_\mathrm{L}(t)$ über zwei Perioden.}
\begin{germanhintblock}
	Der Spannungsabfall über den Transistor ist zu berücksichtigen.
\end{germanhintblock}

% Solution of subtask
\begin{solutionblock}
	The duty cycle corresponds to
	\begin{equation*}
		\begin{aligned}
			D & =\frac{U_\mathrm{out}}{U_\mathrm{in}}=\frac{U_\mathrm{2}}{U_\mathrm{1}-u_\mathrm{on,CE}} \\
			  & =\frac{\SI{12}{\volt}}{\SI{50}{\volt}-\SI{2}{\volt}}=0.25.
		\end{aligned}
		\label{eq:dutycycle}
	\end{equation*}
	If the transistor conducts, the voltage $u_\mathrm{s}$ results in
	\begin{equation*}
		u_\mathrm{s,on}=U_\mathrm{1}-u_\mathrm{on,CE}=\SI{50}{\volt}-\SI{2}{\volt}=\SI{48}{\volt}.
		\label{eq:u_on}
	\end{equation*}

	If the transistor does not conduct, the diode conducts and $u_\mathrm{s}$ yielding
	\begin{equation*}
		u_\mathrm{s,off}=\SI{0}{\volt}.
		\label{eq:u_off}
	\end{equation*}

	This leads to the result
	\begin{equation*}
		u_\mathrm{L}= u_\mathrm{s}-U_\mathrm{2}
		\begin{cases}
			u_\mathrm{L,on}= \SI{48}{\volt}-\SI{12}{\volt} = \SI{36}{\volt}, \quad & \text{if $T_1$ conducts},         \\
			u_\mathrm{L,off}= -\SI{12}{\volt}, \quad                               & \text{if $T_1$ does not conduct}.
		\end{cases}
	\end{equation*}

	Using the voltage at the inductor while the transistor conducts, the current ripple yields:
	\begin{equation*}
		i_\mathrm{L,ripple}= \frac{u_\mathrm{L,on}}{L} \cdot T_\mathrm{on}
		=\frac{u_\mathrm{L,on} \cdot D}{L f_\mathrm{s}}
		=\frac{ \SI{36}{\volt} \cdot 0.25} {\SI{20}{\micro\henry}
			\cdot \SI{100}{\kilo\hertz}} = \SI{4.5}{\ampere}.
	\end{equation*}
	The maximum and minimum inductor current result in:
	\begin{equation*}
		\begin{aligned}
			i_\mathrm{L,max} & = I_\mathrm{2} + 0.5 \cdot i_\mathrm{L,ripple} = \SI{70}{\ampere} + \SI{2.25}{\ampere} = \SI{72.25}{\ampere}  \\
			i_\mathrm{L,min} & = I_\mathrm{2} - 0.5 \cdot i_\mathrm{L,ripple} = \SI{70}{\ampere} - \SI{2.25}{\ampere} = \SI{67.75}{\ampere}.
		\end{aligned}
	\end{equation*}

	% solutionfigure
	\input{fig/Task01/sFigTask1}

\end{solutionblock}

\subtask{ How large is the power demand of the load, if the step-down converter operates
	in boundary conduction mode (BCM)?}{2}

\subtaskGerman{Wie groß ist die aufgenommene Leistung der Last, wenn der Tiefsetzsteller
	im Lückgrenzbetrieb arbeitet?}

% Solution of subtask
\begin{solutionblock}
	The current at boundary conduction mode depends on the current ripple according
	\begin{equation*}
		I_\mathrm{2,BCM}=\frac{i_\mathrm{L,ripple}}{2} =\frac{\SI{4.5}{\ampere}}{2}=\SI{2.25}{\ampere}.
		\label{eq:i_bcm}
	\end{equation*}
	Using the current at boundary conduction mode, the power of the load is obtained by
	\begin{equation*}
		P_\mathrm{load,BCM}=U_\mathrm{2} \cdot I_\mathrm{2}=\SI{12}{\volt} \cdot \SI{2.25}{\ampere}=\SI{27}{\watt}.
		\label{eq:p_bcm}
	\end{equation*}
\end{solutionblock}

\subtask{ In which case the step-down converter operates in discontinous conduction mode (DCM) and what is the effect
	and the potential risk of this mode?}{1}%
\subtaskGerman{In welchem Fall arbeitet der Tiefsetzsteller im Lückbetrieb
	und welche Auswirkungen sowie potenziellen Risiken hat dieser Betriebsmodus?}

% Solution of subtask
\begin{solutionblock}
	If the power demand is less than $P_\mathrm{load,BCM}$ the output voltage increases.
	The step-down converter operated in DCM-mode, if the power consumption is less than $P_\mathrm{load,BCM}$. In this case the current through the inductor is $\SI{0}{\ampere}$ for a certain time at the end of a switching period.
	Risk: compared to CCM, this leads to a voltage increase, which could lead to the damage of overvoltage sensitive components.
\end{solutionblock}

\subtask{Calculate the switching power loss and the total power loss of the IGBT.}{1}%

\subtaskGerman{Berechnen Sie die Schaltverluste sowie die Gesamtleistungsverluste des IGBTs.}

% Solution of subtask
\begin{solutionblock}
	The power loss of the transistor consists of the power loss while conducting the current and the switching losses.
	\begin{equation*}
		P_\mathrm{T_1}=u_\mathrm{on,CE} I_\mathrm{2} D + \left( E_\mathrm{on}+E_\mathrm{off} \right) \cdot f_\mathrm{s}
		= \SI{2}{\volt} \cdot \SI{70}{\ampere} \cdot 0.25 + \left( \SI{20}{\micro\joule} + \SI{40}{\micro\joule} \right)
		\cdot \SI{100}{\kilo\hertz} = \SI{41}{\watt}.
		\label{eq:p_IGBT}
	\end{equation*}
\end{solutionblock}

\subtask{Calculate the efficiency $\eta$ of the step-down converter.}{2}

\subtaskGerman{Berechnen Sie den Wirkungsgrad $\eta$ des Tiefsetzstellers.}

% Solution of subtask
\begin{solutionblock}
	The efficiency is calculated by the power demand of the load divided by the total input power.
	The power demand of the load is
	\begin{equation*}
		P_\mathrm{load}= U_\mathrm{2} \cdot I_\mathrm{2} = \SI{12}{\volt} \cdot \SI{70}{\ampere} = \SI{840}{\watt}.
		\label{eq:P_load}
	\end{equation*}
	Taking into account the power demand, the efficiency is calculated by:
	\begin{equation*}
		\eta=\frac{P_\mathrm{load}}{P_\mathrm{total}} = \frac{P_\mathrm{load}}{P_\mathrm{load} + P_\mathrm{T_1}}
		= \frac{\SI{840}{\watt}}{\SI{840}{\watt} + \SI{41}{\watt}}= 0.95.
		\label{eq:eta}
	\end{equation*}
\end{solutionblock}
