%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Begin exercise %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\ex{Isolated DC-DC converters}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Task 1: Flyback converter %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\task{Flyback converter}
A flyback converter with an input voltage range $U_\mathrm{1} = \SI{300}{\volt} \, \dots \, \SI{900}{\volt}$ is used to supply a control electronics unit. The converter delivers a rated output power of  $P_\mathrm{2} = \SI{30}{\watt}$ at a regulated (constant) output voltage of  $U_\mathrm{2} = \SI{15}{\volt}$. The flyback converter is operated in discontinuous conduction  mode with a constant switching frequency of  $f_\mathrm{s} = \SI{50}{\kilo\hertz}$. The turns ratio of the transformer is $N_\mathrm{1}/N_\mathrm{2}=60/12$, the magnetizing inductance on the primary side is $L_\mathrm{m} = \SI{760}{\micro\henry}$. The coupling between the primary and secondary windings is ideal and the converter operates in steady state.

\input{./fig/ex04/Fig_FlybackConverter.tex}

\begin{table}[ht]
    \centering  % Zentriert die Tabelle
    \begin{tabular}{llll}
        \toprule
        
        Input voltage: &  $U_{\mathrm{1}} = \SI{300}{\volt} \, \dots \, \SI{900}{\volt}$ & Output voltage: & $U_{\mathrm{2}} = \SI{15}{\volt}$ \\ 
        Output power: & $P_2 = \SI{30}{\watt}$  & Transformation ratio: & $N_\mathrm{1}/N_\mathrm{2}=60/12$ \\ 
        Magn. inductance: & $L_\mathrm{m} = \SI{760}{\micro\henry}$ & Switching frequency: & $f_{\mathrm{s}} = \SI{50}{\kilo\hertz}$ \\ 
        \bottomrule
    \end{tabular}
    \caption{Parameters of the flyback converter.}  % Beschriftung der Tabelle
    \label{table:ex04_Parameters of the circuit}
\end{table}

\subtask{The input voltage is $U_\mathrm{1}=\SI{760}{\volt}$ at rated output power. What is the peak value $\hat i_\mathrm{1}$ of the primary current $i_\mathrm{1}$? What is the peak value $\hat i_\mathrm{2}$ of the secondary current $i_\mathrm{2}$? Calculate the duty cycle of the transistor for this operating point.} 

\begin{solutionblock}
To determine the peak current  $\hat i_\mathrm{1}$, we consider the average power transfer
\begin{equation}
    P_\mathrm{2} = W_\mathrm{L} f_\mathrm{s}, \label{eq:output power ex04}
\end{equation}
with the stored magnetic energy
\begin{equation}
    W_\mathrm{L} = \frac{1}{2}L_\mathrm{m}\hat i_\mathrm{1}^2. \label{eq:energy primary inductance ex04}
\end{equation}
Inserting delivers
\begin{equation}
    P_\mathrm{2} = \frac{1}{2}L_\mathrm{1}\hat i_\mathrm{1}^2 f_\mathrm{s}.\label{eq:output power with IDach ex04}
\end{equation}
Solving for $\hat i_\mathrm{1}$ yields the peak current
\begin{equation}
    \hat i_\mathrm{1} = \sqrt{\frac{2P_\mathrm{2}}{L_\mathrm{m}f_\mathrm{s}}}= \sqrt{\frac{2\cdot\SI{30}{\watt}}{\SI{760}{\micro\henry}\cdot\SI{50}{\kilo\hertz}}}=\SI{1.257}{\ampere}.
\end{equation}
Since an ideal transformer is assumed, the secondary peak current directly follows from the primary peak current and the turns ratio:
\begin{equation}
    \hat i_\mathrm{2} = \hat i_\mathrm{1} \frac{N_\mathrm{1}}{N_\mathrm{2}} = \SI{1.257}{\ampere} \cdot \frac{60}{12} = \SI{6.28}{\ampere}.
\end{equation}
Because of DCM the voltage transfer ratio is
\begin{equation}
    \frac{U_2}{U_1} = \frac{D^2}{2} \frac{\Delta i_\mathrm{m,max}}{\overline{i}_2}. \label{eq:Duty cycle ex04}
\end{equation}
The yet unknown current ripple can be determined via the switch-on interval
\begin{equation}
    \Delta i_\mathrm{m,max}= \frac{T_\mathrm{s} \cdot U_1}{L_\mathrm{m}} = \frac{\frac{1}{\SI{50}{\kilo\hertz}}\cdot \SI{760}{\volt}}{\SI{760}{\micro\henry}}=\SI{20}{\ampere}.
\end{equation}
Inserting in \eqref{eq:Duty cycle ex04} and solving for D with $\overline{i}_2 = \frac{P_\mathrm{2}}{U_\mathrm{2}}$ yields to:

\begin{equation}
    D = \sqrt{\frac{2U_2\overline{i}_2}{U_1\Delta i_\mathrm{m,max}}} = \sqrt{\frac{2\cdot \SI{15}{\volt}\cdot\SI{30}{\watt}}{\SI{760}{\volt}\cdot\SI{15}{\volt}\cdot\SI{20}{\ampere}}} = 0.063.
\end{equation}

\end{solutionblock}

\subtask{The input voltage is  $U_\mathrm{1}=\SI{382}{\volt}$ at nominal load. Calculate and sketch the following voltage and current curves for this operating point over two cycle periods: $u_\mathrm{T}(t), u_\mathrm{s}(t), i_\mathrm{2}(t), i_\mathrm{1}(t)$. Here, $u_\mathrm{T}(t)$ is the transistor voltage and $u_\mathrm{s}(t)$ is the voltage on the secondary side of the transformer.}

\begin{solutionblock}
    Because of the different input voltage, the duty cycle $D$ has changed:
    \begin{equation}
        \Delta i_\mathrm{m,max}= \frac{T_\mathrm{s} \cdot U_1}{L_\mathrm{m}} = \frac{\frac{1}{\SI{50}{\kilo\hertz}}\cdot \SI{382}{\volt}}{\SI{760}{\micro\henry}}=\SI{10.05}{\ampere}.
    \end{equation}
    \begin{equation}
        D = \sqrt{\frac{2U_2\overline{i}_2}{U_1\Delta i_\mathrm{m,max}}} = \sqrt{\frac{2\cdot \SI{15}{\volt}\cdot\SI{30}{\watt}}{\SI{382}{\volt}\cdot\SI{15}{\volt}\cdot\SI{10.05}{\ampere}}} = 0.125.
    \end{equation}
    Based on this result, $T_\mathrm{on}$ can be calculated as:
    \begin{equation}
        T_\mathrm{on} = D T_\mathrm{s} = 0.125 \cdot \frac{1}{\SI{50}{\kilo\hertz}} = \SI {2.5}{\micro\s}.
    \end{equation}
    The flyback converter in DCM has three different switch states $T_\mathrm{on}, T'_\mathrm{off}$ and $T''_\mathrm{off}$. Because of this $D'$ has be calculated as:
    \begin{equation}
        D' = \frac{N_\mathrm{2}U_\mathrm{1}}{N_\mathrm{1}U_\mathrm{2}}D = \frac{12\cdot \SI{382}{\volt}}{60 \cdot \SI{15}{\volt}} D = 0.637,
    \end{equation}
    \begin{equation}
        T'_\mathrm{off} = D' T_\mathrm{s} = 0.637 \cdot \frac{1}{\SI{50}{\kilo\hertz}} = \SI {12.7}{\micro\s},
    \end{equation}
    \begin{equation}
        T''_\mathrm{off}= T_\mathrm{s} -T_\mathrm{on}-T'_\mathrm{off} = \frac{1}{\SI{50}{\kilo\hertz}} - \SI {2.5}{\micro\s} - \SI {12.7}{\micro\s} = \SI {4.8}{\micro\s}.
    \end{equation}
    In the following the voltage values in the three intervals are calculated:
    The first interval is $0 < t < T_\mathrm{on}$. In this interval the transistor conducts and, therefore, the voltage $U_\mathrm{T}=0$ results.
    The following applies to the voltage on the primary side:
    \begin{equation}
        u_\mathrm{T} = U_1 + \frac{N_1}{N_2}U_\mathrm{2}.\label{eq:voltageTransistorTask1 ex04}
    \end{equation}
    In this interval the voltage $U_\mathrm{2}$ is equal to $u_\mathrm{s}$. With the voltage $u_\mathrm{T}=0$ follows:
    \begin{equation}
        u_\mathrm{s} = -U_1 \frac{N_2}{N_1} = -\SI{382}{\volt}\cdot\frac{12}{60} = \SI{-76.4}{\volt}.
    \end{equation}
    The second interval is $T_\mathrm{on} < t < T'_\mathrm{off}$. In this interval the transistor blocks. For this calculation \eqref{eq:voltageTransistorTask1 ex04} can be used again as:
    \begin{equation}
        u_\mathrm{T} = \SI{382}{\volt} + \frac{60}{12}\cdot\SI{15}{\volt} = \SI{457}{\volt}.
    \end{equation}
    Because of the conducting diode, for the voltage $U_\mathrm{2}$ is following $U_\mathrm{2} = u_\mathrm{s} = \SI{15}{\volt}$.
    The third interval is $T_\mathrm{on}+T'_\mathrm{off} < t < T_\mathrm{s}$. In this interval the diode is not conducting and the transistor blocks. From this follows $u_\mathrm{T} = U_1$ and $u_\mathrm{s}=0$.
    Looking at the current $i_\mathrm{1}(t)$ over the entire course, the current can only be present when the transistor is switched on. When the transistor is switched on, the current value rises to the peak value $\hat i_\mathrm{1}= \SI{1.257}{\ampere}$. Current $i_\mathrm{2}(t)$ is the discharge current of the secondary winding when the diode is conducting and the transistor blocks. The discharge current starts at the peak value of the secondary side $\hat i_\mathrm{2}= \SI{6.28}{\ampere}$.
    \input{./fig/ex04/Fig_voltageTransistorUsPeriodTask1.tex}
    \input{./fig/ex04/Fig_currentI1I2PeriodTask1.tex}
\end{solutionblock}


\subtask{Determine the mean value $\overline i_\mathrm{T}$ and the RMS current $I_\mathrm{T}$ through the transistor. Also,  determine the mean value $\overline i_\mathrm{D}$ and the RMS current $I_\mathrm{D}$ through the diode. What is the maximum reverse voltage $u_\mathrm{T, max}$ of the transistor and $u_\mathrm{D, max}$ of the diode? Consider the same operation conditions as in the previous subtask.}
\begin{solutionblock}
    The current signal is in the form of a triangle. Therefore, the average value must first be determined over the $T_\mathrm{on}$ interval only. The value is divided by $T_\mathrm{s}$ to obtain the average value over the entire period. This gives:
    \begin{equation}
        \overline{i}_\mathrm{T} = \frac{1}{T_\mathrm{s}}\frac{1}{2}\hat i_\mathrm{1}T_\mathrm{on}=\frac{1}{\SI{20}{\micro\s}}\cdot\frac{1}{2}\cdot\SI{1.257}{\ampere}\cdot\SI{2.5}{\micro\s}=\SI{78.53}{\milli\ampere}.
    \end{equation}
    To determine the RMS value of the transistor current, the function of the transistor current must be set up and used, which leads to
    \begin{equation}
        I_\mathrm{T}^2=\frac{1}{T_\mathrm{s}} \int_{0}^{T_\mathrm{s}} i_\mathrm{1}^2(t) \,\mathrm{d}t \ = \frac{1}{T_\mathrm{s}}\int_{0}^{T_\mathrm{on}} \frac{\hat i_\mathrm{1}^2t^2}{T_\mathrm{on}^2} \,\mathrm{d}t \ =  \frac{\hat i_\mathrm{1}^2T_\mathrm{on}}{3T_\mathrm{on}},
    \end{equation}
    \begin{equation}
        I_\mathrm{T} = \hat i_\mathrm{1} \sqrt{\frac{T_\mathrm{on}}{3T_\mathrm{s}}}= \SI{1.257}{\ampere}\cdot\sqrt{\frac{\SI{2.5}{\micro\s}}{3\cdot\SI{20}{\micro\s}}}= \SI{256.58}{\milli\ampere}.
    \end{equation}
    The mean value for the current $i_\mathrm{D}$ and the RMS value for  $i_\mathrm{D}$ are determined analogously:
    \begin{equation}
        \overline{i}_\mathrm{D} = \frac{1}{T_\mathrm{s}}\frac{1}{2}\hat i_\mathrm{2}T_\mathrm{off}'=\frac{1}{\SI{20}{\micro\s}}\cdot\frac{1}{2}\cdot\SI{6.28}{\ampere}\cdot\SI{12.7}{\micro\s}=\SI{2}{\ampere},
    \end{equation}

    \begin{equation}
        I_\mathrm{D} = \hat i_\mathrm{2} \sqrt{\frac{T_\mathrm{off}'}{3T_\mathrm{s}}}= \SI{6.28}{\ampere}\cdot\sqrt{\frac{\SI{12.7}{\micro\s}}{3\cdot\SI{20}{\micro\s}}}= \SI{2.89}{\ampere}.
    \end{equation}
    The voltage  $u_\mathrm{T,max}$ is calculated as in \eqref{eq:voltageTransistorTask1 ex04} as:
    \begin{equation}
        u_\mathrm{T,max} = U_\mathrm{1} + \frac{N_\mathrm{1}}{N_\mathrm{2}}U_\mathrm{2}= \SI{382}{\volt}+\frac{60}{12}\cdot\SI{15}{\volt}= \SI{457}{\volt}.
    \end{equation}
    \eqref{eq:voltageTransistorTask1 ex04} is used again, but now for the secondary side. This leads to:
    \begin{equation}
        u_\mathrm{D,max} = U_\mathrm{2} + \frac{N_\mathrm{2}}{N_\mathrm{1}}U_\mathrm{1}= \SI{15}{\volt}+\frac{12}{60}\cdot\SI{382}{\volt}= \SI{91.4}{\volt}.
    \end{equation}

\end{solutionblock}

\subtask{How much energy is transferred from the input to the output per switching period $\Delta E$ and what is the resulting average power $P$ (consider the same operation conditions as in the previous subtask)? What happens if there is no ideal voltage source on the output side but an unloaded capacitor and the circuit is operated with $D>0$?}

\begin{solutionblock}
    The energy on the primary side can be determined with \eqref{eq:energy primary inductance ex04}:
    \begin{equation}
        W_\mathrm{L} = \frac{1}{2}\cdot \SI{760}{\micro\henry} \cdot \SI{1.257}{\ampere} = \SI{600}{\micro\joule}.
    \end{equation}
    Since this is an ideal converter, no losses are assumed. 
    To determine the average power ${P}$, the transmitted energy must be divided by the period duration $T_\mathrm{s}$, as follows:
    \begin{equation}
        {P} = \frac{W_\mathrm{L}}{T_\mathrm{s}} = \frac{\SI{600}{\micro\joule}}{\SI{20}{\micro\s}}=\SI{30}{\watt}.
    \end{equation}
At each switching interval, energy is pushed into the capacitor, the voltage of which continues to rise until a component fails due to over voltage if no load is connected.
\end{solutionblock}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Task 2: Forward converter with asymmetric half-bridge
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\task{Forward converter with asymmetric half-bridge}

The schematic of a forward converter with an asymmetric half-bridge is shown in \autoref{fig:ex04_ForwardConverterWithAsymHalfBridge}. 
For the calculations the diodes and transistors are considered as ideal components.

\input{./fig/ex04/Fig_ForwardConverterWithAsymHalfBridge}

The parameters are listed in \autoref{fig:ex04_ForwardConverterWithAsymHalfBridge}.

\input{./fig/ex04/FigTab_ForwardConverterWithAsymHalfBridge}
\FloatBarrier
The leakage inductance, the resistive losses, and the core losses of the transformer are negligible. 
The converter operates in steady-state conditions. Both transistors are controlled by the same signal.

\subtask{At what duty cycle $D$ does the circuit operate?}
\begin{solutionblock}
    The forward converter topology is derived from the buck-topology. Taking into account the turns ratio, the result is
    \begin{equation}
        \frac{U_\mathrm{2}}{U_\mathrm{1}}=D\frac{N_\mathrm{2}}{N_\mathrm{1}}.
        \label{eq:ex04voltageratioforwardconverter}
    \end{equation}
    This leads to
    \begin{equation}
        D=\frac{N_\mathrm{1}}{N_\mathrm{2}}\frac{U_\mathrm{2}}{U_\mathrm{1}}=10\frac{\SI{15}{\volt}}{\SI{325}{\volt}}=0.4615.
    \end{equation}
\end{solutionblock}

\subtask{Calculate the average currents $\overline{i}_\mathrm{2}$ and $\overline{i}_\mathrm{1}$ over a switching cycle assuming ideal filtering of $i_\mathrm{2}$.}
\begin{solutionblock}
    The current is calculated with help of the output power by
    \begin{equation}
        \overline{i}_\mathrm{2}=\frac{P_\mathrm{2}}{U_\mathrm{2}}=\frac{\SI{50}{\watt}}{\SI{15}{\volt}}=\SI{3.33}{\ampere}.
        \label{eq:ex04averageoutputcurrent}
    \end{equation}
    Since the losses can be neglected, it follows
    \begin{equation}
        \overline{i}_\mathrm{1}=\frac{P_\mathrm{2}}{U_\mathrm{1}}=\frac{\SI{50}{\watt}}{\SI{325}{\volt}}=\SI{0.138}{\ampere}.
    \end{equation}
\end{solutionblock}


\subtask{Calculate the peak value $\hat{i}_\mathrm{m}$ of the magnetizing current $i_\mathrm{m}$.}
\begin{solutionblock}
    The duty cycle is less than $0.5$, i.e., the magnetizing current $i_\mathrm{m}$ increase 
    while the transistors are active and decrease to $\SI{0}{\ampere}$ before the next period starts.
    This leads to 
    \begin{equation}
        \hat{i}_\mathrm{m}=\Delta i_\mathrm{m,max}= \frac{D \cdot U_\mathrm{1}}{L_\mathrm{m} \cdot f_\mathrm{s}} =
        \frac{0.4615 \cdot \SI{325}{\volt}}{\SI{2}{\milli\henry} \cdot \SI{50}{\kilo\hertz}} = \SI{1.5}{\ampere}.
    \end{equation}    
\end{solutionblock}


\subtask{Sketch the signals $u_\mathrm{p}$, $i_\mathrm{m}$, $i_\mathrm{p}$ and $i_\mathrm{1}$ 
         considering the switching-induced ripples.}
\begin{solutionblock}
    The current $i_\mathrm{p}$ corresponds to the sum of $i_\mathrm{m}$ and the current $i_\mathrm{1,trOn}$ for the transferred energy to the load.
    Since the losses can be neglected, the portion of the primary current cause by the secondary current is calculated by
    \begin{equation}
        \overline{i}_\mathrm{1sec}=\frac{N_\mathrm{2}}{N_\mathrm{1}}I_\mathrm{2}=\frac{\SI{3.333}{\ampere}}{10}=\SI{0.333}{\ampere}.
    \end{equation}    
    The average value of the current that belongs to the transferred energy, while switch-on phase is calculated by
    \begin{equation}
        \overline{i}_\mathrm{1,trOn}=\frac{\overline{i}_\mathrm{1sec}}{D}=\frac{\SI{0.333}{\ampere}}{0.4615}=\SI{0.722}{\ampere}.
    \end{equation}
    The voltage and currents are displayed in \autoref{fig:ex04_VoltageAtPrimarySide} and \autoref{fig:ex04_CurrentAtPrimarySide}
    \input{./fig/ex04/sFigDia_Ex04SignalsForwardConverterWithAsymHalfBridge}
\end{solutionblock}
\FloatBarrier
\subtask{Calculate the minimal necessary input voltage $U_\mathrm{1}$, if $U_\mathrm{2}$ = \SI{20}{\volt} shall being constant.}
\begin{solutionblock}
    For the forward converter the duty cycle should be less equal $0.5$. Using \eqref{eq:ex04voltageratioforwardconverter} leads to
    \begin{equation}
        U_\mathrm{1}=\frac{N_\mathrm{2}}{N_\mathrm{1}}\frac{U_\mathrm{1}}{D_\mathrm{max}}=
        \frac{\SI{20}{\volt}}{0.5 \cdot 10}=\SI{400}{\volt}.
    \end{equation}
\end{solutionblock}


\subtask{Determine $L$ such that the ripple current $\Delta i_\mathrm{2}$ is $\SI{10}{\percent}$ of the 
         average output current $\overline{i}_\mathrm{2}$.}
\begin{solutionblock}
    The ripple current $\Delta i_\mathrm{2}$ is expressed by
    \begin{equation}
        \Delta i_\mathrm{2}= \frac{D \cdot U_\mathrm{2}}{L_\mathrm{m} \cdot f_\mathrm{s}} = 0.1 \overline{i}_\mathrm{2}.
    \end{equation}
    Using the result of \eqref{eq:ex04averageoutputcurrent} this leads to
    \begin{equation}
        L_\mathrm{m}  = \frac{D \cdot U_\mathrm{2}}{0.1 \overline{i}_\mathrm{2} \cdot f_\mathrm{s}} 
        = \frac{0.4615 \cdot \SI{15}{\volt}}{0.1 \cdot \SI{3.33}{\ampere} \cdot \SI{50}{\kilo\hertz}}=\SI{0.485}{\milli\henry}.
    \end{equation}    
\end{solutionblock}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Task 3: Singled-ended forward converter (demagnetization winding)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\task{Singled-ended forward converter (demagnetization winding)}

The power supply of a data processing system shall be realized by a singled-ended forward converter.

\input{./fig/ex04/Fig_SingledEndedForwardConverter}

\FloatBarrier
The parameters are listed in \autoref{table:Ex04_Parameters of the singled ended forward converter.}.
The output inductance $L$ is dimensioned so that the current $i_\mathrm{L}$ exhibits a continuous waveform.
The transformer's leakage inductance can be neglected.

\input{./fig/ex04/FigTab_SingledEndedForwardConverter}
\vspace{-0.5cm}
\subtask{Calculate the turns ratio $N_\mathrm{3}$/$N_\mathrm{1}$ limiting the maximum transistor blocking voltage 
          to $\SI{600}{\volt}$.}
\begin{solutionblock}
    The maximum blocking voltage at the transistor is calculated by
    \begin{equation}
        U_\mathrm{T,max}=U_\mathrm{1}+U_\mathrm{p} \quad \text{with} \quad U_\mathrm{p}=\frac{N_\mathrm{1}}{N_\mathrm{3}} U_\mathrm{3}.
        \label{eq:ex04transistorblockingvoltageA}
    \end{equation} 
    If the diode $D_\mathrm{3}$ is active the voltage $U_\mathrm{3}$ is equal $U_\mathrm{1}$.
    Using \eqref{eq:ex04transistorblockingvoltageA} leads to
    \begin{equation}
        U_\mathrm{T,max}=U_\mathrm{1}+\frac{N_\mathrm{1}}{N_\mathrm{3}} U_\mathrm{1}.
        \label{eq:ex04transistorblockingvoltageB}
    \end{equation}
    Solving \eqref{eq:ex04transistorblockingvoltageB} to turns ratio $N_\mathrm{3}$/$N_\mathrm{1}$ results in
    \begin{equation}
        \frac{N_\mathrm{3}}{N_\mathrm{1}}=\frac{1}{\frac{U_\mathrm{T,max}}{U_\mathrm{1}}-1}
        =\frac{1}{\frac{\SI{600}{\volt}}{\SI{360}{\volt}}-1}=1.5.
    \end{equation}
\end{solutionblock}
          
\subtask{What is the maximum permissible duty cycle of the power transistor in this case?}
\begin{solutionblock}
    The $\Delta i_\mathrm{m}$ is related to the voltage at the primary coil. The voltages are
    \begin{equation}
        U_\mathrm{p,on}=U_\mathrm{1}  \quad \text{and} \quad U_\mathrm{p,off}=U_\mathrm{1}\frac{N_\mathrm{3}}{N_\mathrm{1}} U_\mathrm{3}.
        \label{eq:ex04upmagnetizationvoltage}
    \end{equation}     
    The magnetization current is $\SI{0}{\ampere}$ at the end of each period. For the maximum duty cycle this leads to
    \begin{equation}
        U_\mathrm{p,on}T_\mathrm{p,on}=U_\mathrm{p,off}T_\mathrm{p,off}.
        \label{eq:ex04upmagnetizationtimerelation}        
    \end{equation} 
    The ration of $T_\mathrm{p,on}/T_\mathrm{p,off}$ is obtained by
    \begin{equation}
        \frac{T_\mathrm{p,on}}{T_\mathrm{p,off}}=\frac{D_\mathrm{max}}{1-D_\mathrm{max}}.
    \end{equation} 
    Using \eqref{eq:ex04upmagnetizationvoltage} and \eqref{eq:ex04upmagnetizationtimerelation} results in
    \begin{equation}
        U_\mathrm{p,on} \frac{D_\mathrm{max}}{1-D_\mathrm{max}} = U_\mathrm{p,off}.
        \label{eq:ex04upmagnetizationdutycycle}
    \end{equation}
    The cancelation of voltages $U_\mathrm{p,on}$ and $U_\mathrm{p,off}$ results in
    \begin{equation}
        \frac{N_\mathrm{3}}{N_\mathrm{1}} \frac{D_\mathrm{max}}{1-D_\mathrm{max}} = 1.
        \label{eq:ex04upmagnetizationdutycycle_sol}
    \end{equation} 

    The duty cycle is calculated by solving \eqref{eq:ex04upmagnetizationdutycycle_sol}:
    \begin{equation}
        D_\mathrm{max}=\frac{\frac{N_\mathrm{1}}{N_\mathrm{3}}}{1+\frac{N_\mathrm{1}}{N_\mathrm{3}}} 
        = \frac{0.667}{1+0.667}=0.4.
    \end{equation}  
\end{solutionblock}

\subtask{What turns ratio $N_\mathrm{1}$/$N_\mathrm{2}$ should be chosen to achieve the required secondary voltage?}
\begin{solutionblock}
    The maximal duty cycle is applied at minimum input voltage. Moreover, for the calculation of the 
    turns ratio $N_\mathrm{1}$/$N_\mathrm{2}$ the forward voltage of $D_\mathrm{1}$ has to taken in account.
    The voltage at the secondary coil is calculated by
    \begin{equation}
        U_\mathrm{2}=\left(U_\mathrm{s}-U_\mathrm{D1,f}\right) D \quad  \mbox{and} \quad 
        U_\mathrm{s}=\frac{N_\mathrm{2}}{N_\mathrm{1}}U_\mathrm{1}.
        \label{eq:ex04transformationratio}        
    \end{equation}
    This leads to
    \begin{equation}
        U_\mathrm{2}=\left(\frac{N_\mathrm{2}}{N_\mathrm{1}}U_\mathrm{1}-U_\mathrm{D1,f}\right) D.
        \label{eq:ex04secondaryvoltage}  
    \end{equation}
    Solving \eqref{eq:ex04secondaryvoltage} with respect to $N_\mathrm{1}$/$N_\mathrm{2}$ results in
    \begin{equation}
        \frac{N_\mathrm{1}}{N_\mathrm{2}}=\frac{U_\mathrm{1}D}{U_\mathrm{2}+ U_\mathrm{D1,f}D}
        =\frac{\SI{240}{\volt}\cdot 0.4}{\SI{5}{\volt}+\SI{0.4}{\volt} \cdot 0.4 }=18.6.
    \end{equation}
\end{solutionblock}



\subtask{Does the duty cycle need to be adjusted when the output power changes? 
         Over what range must the duty cycle be adjustable, considering the input voltage range?}
\begin{solutionblock}
    No, \eqref{eq:ex04transformationratio} shows that the output voltage depends on the turn ratio and the duty cycle,
    but not on the output power. The stored magnetic energy of the transformer is not used for the energy transfer.
    This answer is only valid the with the assumption of ideal components. 
    The minimum duty cycle results from solving \eqref{eq:ex04secondaryvoltage} with respect to D:
    \begin{equation}
        D=\frac{U_\mathrm{2}}{\frac{N_\mathrm{2}}{N_\mathrm{1}}U_\mathrm{1}-U_\mathrm{D1,f}}=
        \frac{\SI{5}{\volt}}{\frac{N_\mathrm{2}}{N_\mathrm{1}}\SI{360}{\volt}-\SI{0.4}{\volt}}=0.264.
    \end{equation}
    The duty cycle needs to be adjustable between $0.264$ and $0.4$.
\end{solutionblock}
         
\subtask{What are the resulting maximum blocking voltages of the diodes $D_\mathrm{1}$ and $D_\mathrm{2}$?}
\begin{solutionblock}
    The maximum blocking voltages of the diodes $D_\mathrm{1}$ is calculated by
    \begin{equation}
        U_\mathrm{D2,r}=U_\mathrm{s}-U_\mathrm{D1,f} \quad \text{with} \quad U_\mathrm{s}=U_\mathrm{1}\frac{N_\mathrm{2}}{N_\mathrm{1}}.
    \end{equation}
    This leads to
    \begin{equation}
        U_\mathrm{D2,r}=U_\mathrm{1}\frac{N_\mathrm{2}}{N_\mathrm{1}}--U_\mathrm{D1,f}=
        \SI{360}{\volt}\frac{1}{18.6}-\SI{0.4}{\volt}=\SI{18.95}{\volt}.
    \end{equation}
    The maximum blocking voltages of the diodes $D_\mathrm{1}$ corresponds to the secondary voltage at demagnetization.
    Therefore, the turn ratio of $N_\mathrm{2}/N_\mathrm{3}$ is used:
    \begin{equation}
        U_\mathrm{D1,r}=U_\mathrm{s}=U_\mathrm{1}\frac{N_\mathrm{2}}{N_\mathrm{3}} \quad \text{with} \quad
        \frac{N_\mathrm{2}}{N_\mathrm{3}}=\frac{N_\mathrm{2}}{N_\mathrm{1}}\frac{N_\mathrm{1}}{N_\mathrm{3}}.
    \end{equation}
    The result is obtained by
    \begin{equation}
        U_\mathrm{D1,r}=\SI{360}{\volt}\frac{1}{18.6 \cdot 1.5}=\SI{12.9}{\volt}.
    \end{equation}
\end{solutionblock}

\subtask{Determine the magnetizing inductance $L_\mathrm{m}$ to ensure 
        that the peak value of the magnetizing current remains below $\SI{10}{\percent}$ of the $\overline{i}_\mathrm{L}'$, which corresponds to the average current $\overline{i}_\mathrm{L}$ through the output inductance translated to the primary side
        at a nominal load of $P_2=\SI{125}{\watt}$.}
\begin{solutionblock}
    The current through $L$ is calculated by
    \begin{equation}
        I_\mathrm{2}=P_\mathrm{2}/U_\mathrm{2}=\frac{\SI{125}{\watt}}{\SI{5}{\volt}}=\SI{25}{\ampere}
        \label{eq:ex04ouputcurrent}.
    \end{equation}
    This corresponds to the current at the primary side of
    \begin{equation}
        I_\mathrm{2s}'=I_\mathrm{2}\frac{N_\mathrm{2}}{N_\mathrm{1}}.
    \end{equation}
    This leads to the current $\hat{i}_\mathrm{m}$ of
    \begin{equation}
        \hat{i}_\mathrm{m}=I_\mathrm{2s}'\cdot 0.1=I_\mathrm{2}\frac{N_\mathrm{2}}{N_\mathrm{1}}\cdot 0.1
        =\frac{\SI{25}{\ampere}}{18.6} \cdot 0.1 =\SI{0.1344}{\ampere}.
        \label{eq:ex04magnetizationcurrent}  
    \end{equation}
    The maximum magnetization current maximum $\hat{i}_\mathrm{m}$ corresponds to the ripple current $\Delta{i}_\mathrm{m}$,
    because at the end of each period the magnetization current $i_\mathrm{m}=\SI{0}{\ampere}$:
    \begin{equation}
        \Delta i_\mathrm{m}= \frac{D \cdot U_\mathrm{1}}{L_\mathrm{m} \cdot f_\mathrm{s}}.
        \label{eq:ex04ripplemagnetizationcurrent}
    \end{equation}
    Using \eqref{eq:ex04magnetizationcurrent} and solving \eqref{eq:ex04ripplemagnetizationcurrent} 
    with respect to $L_\mathrm{m}$ yields
    \begin{equation}
        L_\mathrm{m}= \frac{D_\mathrm{max} \cdot U_\mathrm{1}}{\hat{i}_\mathrm{m} \cdot f_\mathrm{s}}
                    = \frac{0.4 \cdot \SI{240}{\volt}}{\SI{0.1344}{\ampere} \cdot \SI{48}{\kilo\hertz}}=\SI{14.9}{\milli\henry}.
    \end{equation}
\end{solutionblock}

\subtask{Sketch the signals of the voltage across the power transistor, the current through the demagnetization 
        winding, and the current through the freewheeling diode $D_\mathrm{2}$ 
        for $U_\mathrm{1}=\SI{240}{\volt}$ and $U_\mathrm{1}=\SI{360}{\volt}$.}
\begin{solutionblock}
    The current through the demagnetization is calculated by
    \begin{equation}
        \hat{i}_\mathrm{3}=\frac{N_\mathrm{3}}{N_\mathrm{1}} \cdot \Delta i_\mathrm{m}
        =\frac{1}{1.5} \cdot \SI{0.1344}{\ampere}=\SI{89.6}{\milli\ampere}.
    \end{equation}
    The current through freewheeling diode $D_\mathrm{2}$ is taken from \eqref{eq:ex04ouputcurrent}.
    The voltages and currents are displayed in \autoref{fig:ex04_VoltageAtTransistor}, 
    \autoref{fig:ex04_DemagnetizationCurrentN3} and \autoref{fig:ex04_D2FreewheelingCurrent} for the two input voltages.
    \input{./fig/ex04/sFigDia_Ex04SignalsSingledEndedForwardConverter}  
\end{solutionblock}
        
\subtask{Calculate the peak magnetizing current for each case assuming a constant output current.}
\begin{solutionblock}
    The current through the primary coil $N1$ is still calculated with \eqref{eq:ex04magnetizationcurrent} and yields
    $\SI{0.1344}{\ampere}$. 
    Using \eqref{eq:ex04ripplemagnetizationcurrent} the minimal duty cycle yields
    \begin{equation}
        \hat{i}_\mathrm{m,D_{min}}=\Delta i_\mathrm{m,D_{min}}= \frac{D_\mathrm{min} \cdot U_\mathrm{1}}{L_\mathrm{m} \cdot f_\mathrm{s}}
        =\frac{0.4 \cdot \SI{360}{\volt}}{\SI{14.9}{\milli\henry} \cdot \SI{48}{\kilo\hertz}}=\SI{0.1330}{\ampere}.
    \end{equation}
\end{solutionblock}


\subtask{Could a higher power be transferred by doubling the switching frequency of the converter?}
\begin{solutionblock}
    No,  also in this case \eqref{eq:ex04transformationratio} shows, that the output voltage depends on the turn ratio and the duty cycle,
    but not on the switching frequency of the converter. The transformer does not act as energy storage.
    Also, this answer is valid under the assumption of ideal components. 
\end{solutionblock}