%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Begin exercise %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ex{Rectifiers}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Task 1: B2U topology with capacitive filtering %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\task{B2U topology with capactive filtering}
An uncontrolled single-phase, two-pulse rectifier circuit with capacitive filtering 
is shown in \autoref{fig:B2U_Topology_Cap_Filtering}.
All components, including the diodes, are assumed to be ideal. On the input side, 
the single-phase AC supply with voltage $u_\mathrm{1}(t)$ is connected, while on the output side, 
a smoothing capacitor $C$ and a constant current load $I_\mathrm{0}$ are present.

\input{fig/ex05/Fig_B2U_TopologyWithCapactiveFiltering.tex}

\begin{table}[ht]
    \centering  % Zentriert die Tabelle
    \begin{tabular}{llll}
        \toprule
        
        Input voltage: &  $u_\mathrm{1}(t) = \SI{156}{\volt}\cdot \sin(\omega t)$ & Load current: & $I_{\mathrm{0}} = \SI{7.5}{\ampere}$ \\ 
        Filter capacitance: & $C = \SI{330}{\micro\farad}$  & Frequency: & $f= \SI{60}{\hertz}$ \\ 
        \bottomrule
    \end{tabular}
    \caption{Parameters of the B2U rectifier.}  
    \label{table:ex05_Task1_ParametersOfTheCircuit}
\end{table}

The angle $\alpha$ represents the phase angle range between zero crossing of the supply voltage and the phase angle
at which all four diodes are blocked, meaning the capacitor discharges through the load. The angle $\beta$ 
represents the phase angle range between $\alpha$ and the phase angle at which two of the four diodes begin to 
conduct, i.e., when the capacitor is recharged from the mains supply.
A steady-state operation is assumed for this task.

\subtask{Calculate the two angles $\alpha$ and $\beta$.
Note: For the calculation of $\beta$ you can use the following simple approximation: $\sin(x) \approx x$.
(This approximation is sufficiently accurate within a range of approximately $x=\pm \SI{25}{\degree}$.)}
\begin{solutionblock}
    As long as the current $i_\mathrm{2}(t)>\SI{0}{\ampere}$ the voltage at the 
    capacitor follows the absolute value of $u_\mathrm{1}(t)$:
    \begin{equation} 
        u_\mathrm{2}(t) = \left| \hat{u}_\mathrm{1}\sin(\omega t)\right| \quad \text{for} \quad i_\mathrm{2}(t)>\SI{0}{\ampere}.
        \label{eq:u2_conductionphase}
    \end{equation}
    While the conduction phase $(i_\mathrm{2} > \SI{0}{\ampere})$ the capacitor is charged, otherwise the capacitor provides the load current. 
    The current of the capacitor $i_\mathrm{C}$ is expressed by
    \begin{equation}
        i_\mathrm{C}(t) = \begin{cases}
            -I_\mathrm{0}, & i_2(t) = 0,\\
            C\frac{\mathrm{d}}{\mathrm{d}t}u_2(t), & i_2(t) > 0,
        \end{cases}.
        \label{eq:iC_B2U_cap_filt}
    \end{equation}
    In the conduction phase $i_\mathrm{2}(t)$ is calculated by:
    \begin{equation}
        i_2(t) = i_\mathrm{C}(t) + I_\mathrm{0}.
        \label{eq:i2_B2U_cap_filt_gen}
    \end{equation}
    Using \eqref{eq:u2_conductionphase} and \eqref{eq:iC_B2U_cap_filt} in \eqref{eq:i2_B2U_cap_filt_gen} results in
    \begin{equation}
        i_2(t) = C\frac{\mathrm{d}}{\mathrm{d}t}\left| \hat{u}_\mathrm{1}\sin(\omega t)\right| + I_0=
        C\omega \hat{u}_1 \cos(\omega t) + I_\mathrm{0}, \quad \text{for} \quad 0 \leq \omega t < \omega t_1.
        \label{eq:i2_B2U_cap_filt}
    \end{equation}
    In \eqref{eq:i2_B2U_cap_filt} the angle $\omega t_1$ corresponds to the angle $\alpha$ at which the conduction phase ends due to
    $i_\mathrm{2}(t_1)=\SI{0}{\ampere}$. Enter $t_1$ in \eqref{eq:i2_B2U_cap_filt} leads to:
    \begin{equation}
        0 = C\omega \hat{u}_\mathrm{1} \cos(\omega t_\mathrm{1}) + I_\mathrm{0} = C\omega \hat{u}_1 \cos(\alpha) + I_\mathrm{0} .
        \label{eq:alpha_B2U_cap_filt_gen}
    \end{equation}    
    Solving \eqref{eq:alpha_B2U_cap_filt_gen} with respect to $\alpha$ results in
    \begin{equation}
        \alpha = \arccos\left(-\frac{I_\mathrm{0}}{C\omega \hat{u}_1}\right) = 
        \arccos(-\frac{\SI{7.5}{\ampere}}{2 \pi \cdot \SI{60}{\hertz} \cdot \SI{330}{\micro\farad} \cdot \SI{156}{\volt}}) = \SI{112.8}{\degree}.
        \label{eq:alpha_B2U_cap_filt_res}
    \end{equation}    
    Starting at angle $\alpha$ the capacitor is discharging by the load current $I_\mathrm{0}$ up to the point 
    $\omega t_2$, where $\left| u_\mathrm{1}(t_2) \right|=u_\mathrm{2}(t_2)$.
    In the phase angle range $\omega t_1 \leq \omega t < \omega t_2$ the voltage $u_\mathrm{2}(t)$ is:
    \begin{equation}
        u_\mathrm{2}(t) = u_2(\omega t_1) + \int_{t_1}^t -\frac{I_0}{C} \mathrm{d} \tau 
        = u_\mathrm{2}(\alpha) + \int_{\alpha}^{\omega t} -\frac{I_0}{\omega C} \mathrm{d} \omega \tau
        =  u_\mathrm{2}(\alpha) - \frac{I_\mathrm{0}}{\omega C} (\omega t - \alpha).
        \label{eq:u2_blockingphase}
    \end{equation}
    Using \eqref{eq:u2_blockingphase} with the condition $\left| u_\mathrm{1}(t_\mathrm{2}) \right|=u_\mathrm{2}(t_\mathrm{2})$ leads to
    \begin{equation}
        u_\mathrm{2}(\alpha) - \frac{I_\mathrm{0}}{\omega C} (\omega t_\mathrm{2} - \alpha) = \left| \hat{u}_\mathrm{1}\sin(\omega t_\mathrm{2})\right|.
        \label{eq:alphabeta__B2U_cap_filt_gen}
    \end{equation}
    This is solvable for $\omega t_\mathrm{2}$ in closed-form by using the approximation $\sin(\omega t) \approx \omega t$. The voltage 
    $\left| u_\mathrm{1}(\omega t)\right|$ starts increasing at $\omega t > \pi$ with $\hat{u}_\mathrm{1} sin(\omega t - \pi)$.
    Using  \eqref{eq:alphabeta__B2U_cap_filt_gen} we obtain
    \begin{equation}
        u_2(\alpha) - \frac{I_\mathrm{0}}{\omega C} (\omega t_\mathrm{2} - \alpha) = \hat{u}_\mathrm{1}(\omega t_\mathrm{2} - \pi).
        \label{eq:alphabeta__B2U_cap_filt_res}
    \end{equation}
    Solving \eqref{eq:alphabeta__B2U_cap_filt_res} with respect to $\omega t_\mathrm{2}$ leads to
    \begin{equation}
        \omega t_\mathrm{2} = \frac{u_2(\alpha) + \frac{I_\mathrm{0}\alpha}{\omega C} + \hat{u}_\mathrm{1} \pi}{\hat{u}_\mathrm{1} + \frac{I_\mathrm{0}}{\omega C}}
                             =\frac{\SI{143.4}{\volt} +\frac{\SI{7.5}{\ampere} \cdot \SI{1.969}{\radian}}{\pi \cdot \SI{60}{\hertz} \cdot \SI{330}{\micro\farad}} + \SI{156}{\volt} \cdot \pi}
                             {\SI{156}{\volt} + \frac{\SI{7.5}{\ampere} }{\pi \cdot \SI{60}{\hertz} \cdot \SI{330}{\micro\farad}}}=\SI{3.478}{\radian}=\SI{199.2}{\degree}.
    \end{equation}
    The angle $\beta$ is calculated by
    \begin{equation}
        \omega t_\mathrm{2} = \alpha + \beta \quad \Rightarrow \quad \beta = \omega t_\mathrm{2} - \alpha = 
        \SI{199.2}{\degree} - \SI{112.8}{\degree} = \SI{86.4}{\degree}.
    \end{equation}
\end{solutionblock}

\vspace{2em}\par

\subtask{Sketch the capacitor voltage $u_\mathrm{2}(\omega t)$ considering $\omega t \in [0,...,2\pi]$ taking 
into account the previously calculated angles $\alpha$ and $\beta$.}
\begin{solutionblock}
    \autoref{sfig:ex05_Voltage_u2_andCurrent_i1_ic} displays the voltage at the capacitor.
    \input{fig/ex05/sFig_VoltageU2AndCurrenti1_ic}  
\end{solutionblock}

\subtask{Calculate the currents $i_\mathrm{1}(\omega t)$  and $i_\mathrm{C}(\omega t)$ and and add them to the previous plot.}
\begin{solutionblock}
    Two cases are to consider for the currents of $i_\mathrm{1}(\omega t)$ and  $i_\mathrm{C}(\omega t)$. If $\alpha<\omega t<\alpha+\beta$
    and $\pi+\alpha<\omega t<\pi+\alpha+\beta$, the capacitor supplies the load current $I_{\mathrm{0}}$.
    The remaining phase angle range the input voltage supplies the load current $I_{\mathrm{0}}$ and charge the capacitor.
    For $\alpha<\omega t<(\alpha+\beta)$ and $(\pi+\alpha)<\omega t<(\pi+\alpha+\beta)$ the current $i_\mathrm{C}(\omega t)$ 
    corresponds to the current $-I_{\mathrm{0}}$ while the diode bridge blocks:
    \begin{equation} 
        i_\mathrm{C}(\omega t)=-I_{\mathrm{0}} \quad \text{and} \quad i_\mathrm{1}(\omega t)= i_\mathrm{2}(\omega t)= \SI{0}{\ampere}.
    \end{equation}
    For $(\alpha+\beta-\pi)\leq\omega t\leq\alpha$ and $(\alpha+\beta)\leq\omega t\leq(\pi+\alpha)$ the current $i_\mathrm{C}(\omega t)$ results in
    \begin{equation}
        \begin{aligned}
            i_\mathrm{C}(\omega t)&=C\frac{\mathrm{d}(u_\mathrm{2}(t))}{\mathrm{dt}}=C\frac{\mathrm{d}(\left| \hat{u}_\mathrm{1}\sin(\omega t)\right|)}{\mathrm{dt}}
            = \omega C \left|\hat{u}_\mathrm{1}\cos(\omega t)\right|  \\
            &= \SI{377}{\hertz} \cdot \SI{330}{\micro\farad} \cdot  \left|\SI{156}{\volt}\cos(\omega t)\right|
            =\SI{19.4}{\ampere}\left| \cdot \cos(\omega t)\right|.
        \end{aligned}
    \end{equation}
    The current $i_\mathrm{2}(\omega t)$ is the sum of $I_{\mathrm{0}}$ and $i_\mathrm{C}(\omega t)$:
    \begin{equation} 
        i_\mathrm{2}(\omega t)=i_\mathrm{C}(\omega t) + I_{\mathrm{0}}=\SI{19.4}{\ampere}\left|\cos(\omega t)\right| + \SI{7.5}{\ampere}.
    \end{equation}
    The current $i_\mathrm{1}(\omega t)$ depends on the diodes, which conduct.
    \begin{equation}
        \begin{split}
            i_\mathrm{1}(\omega t)=i_\mathrm{2}(\omega t) \quad \text{for} \quad (\alpha+\beta-\pi)\leq\omega t\leq\alpha \\
            i_\mathrm{1}(\omega t)=-i_\mathrm{2}(\omega t) \quad \text{for} \quad (\alpha+\beta)\leq\omega t\leq\pi+\alpha.
        \end{split}
    \end{equation}

    The currents $i_\mathrm{C}(\omega t)$ and $i_\mathrm{1}(\omega t)$ added to \autoref{sfig:ex05_Voltage_u2_andCurrent_i1_ic}.

\end{solutionblock}

\subtask{Assume the smoothing capacitor is very large, i.e., $C\rightarrow \infty$. What is the average active power $P_\mathrm{0}$ 
absorbed by the current source? What will $P_\mathrm{0}$ be if $C=\SI{330}{\micro\farad}$?}
\begin{solutionblock}
    Considering \eqref{eq:alpha_B2U_cap_filt_res} the term $I_\mathrm{0}/{\omega C}$ becomes zero, if $C\rightarrow \infty$. 
    This results in $\alpha=\SI{90}{\degree}$, which leads to
    \begin{equation} 
        u_\mathrm{2}(\SI{90}{\degree})=\hat{u}_\mathrm{1}(\sin(\SI{90}{\degree}))=\hat{u}_\mathrm{1}.
    \end{equation}
    For a capacitor with $C\rightarrow \infty$ the voltage decrease is zero, so that the voltage $u_\mathrm{2}(t)$ is constant.
    In this case the power absorbed by the load current is obtained by
    \begin{equation} 
        P_\mathrm{0}=\hat{u}_\mathrm{1} \cdot I_{\mathrm{0}}=\SI{156}{\volt}\cdot\SI{7.5}{\ampere}=\SI{1167}{\watt}.
    \end{equation}
    For the capacitor with \SI{330}{\micro\farad} again the two phase angle ranges are to distinguish. First, the phase angle range
    $(\alpha+\beta-\pi)\leq\omega t\leq\alpha$ is considered. The start angle of the conduction phase is calculated by
    \begin{equation} 
        \beta'=\alpha+\beta-\pi=\SI{1.969}{\radian}+\SI{1.51}{\radian}-\pi
        =\SI{0.337}{\radian} = \SI{19.3}{\degree}.
    \end{equation}
    Using $\beta'$ the absorbed average power of the load current is calculated by:
    \begin{equation} 
        P_\mathrm{0,1}=\frac{\hat{u}_\mathrm{1} \cdot I_{\mathrm{0}} \cdot \omega}{\alpha-\beta'} \int_{\beta'}^{\alpha} \sin(\omega t) dt.
    \end{equation}
    This leads to
    \begin{equation} 
        \begin{aligned}
            P_\mathrm{0,1}&=\frac{\hat{u}_\mathrm{1} \cdot I_{\mathrm{0}}}{\alpha-\beta'} \left( \cos(\beta') - \cos(\alpha) \right) \\
            &=\frac{\SI{156}{\volt} \cdot \SI{7.5}{\ampere}}{\SI{1.969}{\radian}-\SI{0.337}{\radian}}
             \left( \cos(\SI{0.337}{\radian}) - \cos(\SI{1.969}{\radian}) \right)=\SI{952}{\watt}.
        \end{aligned}
    \end{equation}
    For the phase angle range $\alpha<\omega t<(\alpha+\beta)$ the voltage $u_\mathrm{2}(\omega t)$ is linear and the current is independent of the voltage.
    This means that the power is a linear function of the voltage. The average power is obtained by the load current multiplied
    with the average voltage:
    \begin{equation} 
        P_\mathrm{0,2}=I_{\mathrm{0}} \cdot \hat{u}_\mathrm{1}\frac{\sin(\alpha) + \sin(\beta')}{2}=
        \SI{7.5}{\ampere} \cdot \SI{156}{\volt} \cdot \frac{\sin(\SI{112.8}{\degree}) + \sin(\SI{19.3}{\degree})}{2}=
        \SI{733}{\watt}.
    \end{equation}
    The total absorbed average power of the load current results from the weighted sum of $P_\mathrm{0,1}$ and $P_\mathrm{0,2}$:
    \begin{equation} 
        P_\mathrm{0}=\frac{(\SI{180}{\degree}-\beta) \cdot P_\mathrm{0,1}+ \beta \cdot P_\mathrm{0,2}}{\SI{180}{\degree}}
        =\frac{(\SI{180}{\degree}-\SI{86.5}{\degree}) \cdot \SI{952}{\watt}+\SI{86.5}{\degree} \cdot \SI{733}{\watt}}{\SI{180}{\degree}}=\SI{847}{\watt}.
    \end{equation}
    
\end{solutionblock}

\subtask{What is the minimum blocking voltage ratings of the diodes to ensure that the rectifiers is not damaged?}
\begin{solutionblock}
    If voltage $\left| u_\mathrm{1}(t) \right| < u_\mathrm{2}(t)$ the diode bridge blocks and the connection points of $u_\mathrm{1}(t)$
    floats within the voltage range of $u_\mathrm{2}(t)$. The minimum blocking voltage ratings is to calculate, if a diode pair conducts.
    The positive maximum of ${u}_\mathrm{1}(t)=\hat{u}_\mathrm{1}=\SI{156}{\volt}$. In this case $D_\mathrm{1}$
    and $D_\mathrm{2}$ conduct and the voltage drops at $D_\mathrm{3}$ and $D_\mathrm{4}$ in block direction are $\hat{u}_\mathrm{1}=\SI{156}{\volt}$.
    The negative maximum of ${u}_\mathrm{1}(t)=-\hat{u}_\mathrm{1}=-\SI{156}{\volt}$. In this case $D_\mathrm{3}$
    and $D_\mathrm{4}$ conduct and the voltage drops at $D_\mathrm{1}$ and $D_\mathrm{2}$ in block direction are $\hat{u}_\mathrm{1}=\SI{156}{\volt}$.
    This leads to the result, that the minimum blocking voltage ratings of all diodes yields $\SI{156}{\volt}$.
\end{solutionblock}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Task 2: PFC rectifier %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\task{PFC rectifier}
Due to the constantly increasing load on the grid with harmonics as a result of the use of power converters, the regulations regarding the permissible harmonic content of the current consumption of electrical consumers are being tightened. It is therefore necessary, e.g. for the rectification of single-phase AC mains voltage, to design power converters with a high power factor. 
A variant of a PFC rectifier circuit is shown in \autoref{fig:Boost converter with single-phase diode bridge_topology}. The prerequisite for the use of the boost converter is: $u_\mathrm{2} = U_\mathrm{2}>u'(t)$. The boost converter is operated with a pulse width modulated (PWM)-based controller for which the switching frequency $f_\mathrm{T}$ has a constant value $f_\mathrm{T} = \SI{20}{\kilo\hertz}$. CCM is assumed as the operating mode.
\input{fig/ex05/Fig_BoostConverter_with_Rectifiers.tex}

\begin{table}[ht]
    \centering  % Zentriert die Tabelle
    \begin{tabular}{llll}
        \toprule
        
        Input voltage: &  $u_{\mathrm{1}}(t) = \hat u_{\mathrm{1}} \sin(\omega t) = \sqrt{2} \cdot \SI{230}{\volt} \cdot \sin(\omega t)$ & Output voltage: & $u_{\mathrm{2}}(t) = \SI{400}{\volt}$ \\ 
        Output power: & $P_\mathrm{2} = \SI{4}{\kilo\watt}$  & Grid frequency: & $ f =  \SI{50}{\hertz}$ \\ 
        Inductance: & $L = \SI{570}{\micro\henry}$
         & Switching frequency: & $f_\mathrm{s} = \SI{20}{\kilo\hertz}$\\
        \bottomrule
    \end{tabular}
    \caption{Parameters of the PFC rectifier.}  
    \label{table:ex05_Parameters of the circuit}
\end{table}

\subtask{Specify the voltage transformation ratio $m(t)= \frac{u_{\mathrm{2}}(t)}{u'(t)}$ as a function of the duty cycle $d(t)$.}
\begin{solutionblock}
    The equation for the average voltage $U_{\mathrm{L}}$ during a pulse period is given with:
    \begin{equation}
        U_{\mathrm{L}} = (d(t) u'(t) + (1-d(t))(u'(t)-U_{\mathrm{2}})). 
    \end{equation}
The average voltage over the inductance for a (quasi) steady-state operation is zero $U_{\mathrm{L}} = \SI{0}{\volt}$: 
    \begin{equation}
    0 = (d(t) u'(t) + (1-d(t))(u'(t)-U_{\mathrm{2}}))f_\mathrm{s}. \label{eq:ex05ratio2.2}
    \end{equation}   

    Rewriting \eqref{eq:ex05ratio2.2} delivers  the voltage transformation ratio:
  \begin{equation}
    m(t) = \frac{U_{\mathrm{2}}}{u'(t)}=\frac{1}{1-d(t)}. \label{eq:ex05ratio_m(t)}
  \end{equation}
\end{solutionblock}

\subtask{Specify the conduction time of the transistor and the diode as a function of the transformation ratio $M$ and the time $t$, with the assumption $u_2(t)\approx U_2$.} 
\begin{solutionblock}
    Using $M = \frac{U_{\mathrm{2}}}{\hat u_{\mathrm{1}}}$, the transformation ratio is given with:
    \begin{equation}
        m(t) = \frac{U_{\mathrm{2}}}{ u_{\mathrm{1}}(t)}=\frac{U_{\mathrm{2}}}{\hat u_{\mathrm{1}} \sin(\omega t)} = M \frac{1}{\sin(\omega t)}.
    \end{equation}
   If the transistor $T$ is conducting, the duty cycle $d(t)$ is derived from \eqref{eq:ex05ratio_m(t)} as:
\begin{equation}
    d(t) = 1-\frac{1}{m(t)} = 1- \frac{u_{\mathrm{1}}(t)}{U_{\mathrm{2}}}=1- \frac{\hat u_{\mathrm{1}}\sin(\omega t)}{U_{\mathrm{2}}} = 1 -\frac{1}{M} \sin(\omega t).
\end{equation}
If the diode $D_{\mathrm{5}}$ is conducting, the duty cycle $1-d(t)$ is derived from \eqref{eq:ex05ratio_m(t)} as:
\begin{equation}
    1-d(t) = \frac{1}{m(t)}=\frac{1}{M} \sin(\omega t).
\end{equation}
\end{solutionblock}

\subtask{Calculate the maximum amplitude of the switching-induced ripple of the mains current $i_\mathrm{1}$ for the specified operating point.
Note: Consider the conductive state of $T$ and set the voltage across the inductance as a function of the phase angle $(\omega t)$ and the conduction time of the transistor according to the previous subtask.} 
\input{fig/ex05/Fig_courses_i'}
\begin{solutionblock}
    To solve this task, the component equation of inductance is used:
    \begin{equation}
        u_{\mathrm{L}}(t) = L \frac{\mathrm{d}i'(t)}{\mathrm{d}t}.
    \end{equation}
    As a simplification, it is assumed that the $\Delta$ is an approximation of the total differential by a difference equation with the differences $\Delta$. This leads to the component differential equations becoming mean value equations such as:
    \begin{equation}
        \frac{\Delta i'(t)}{\Delta t} = \frac{U_{\mathrm{L}}}{L}.
    \end{equation}
    When transistor $T$ is conducting it applies $U_{\mathrm{L}} = u_{\mathrm{1}}(t) = \hat u_{\mathrm{1}} \sin(\omega t)$.
    This means for $\Delta i'(t)$:
\begin{equation}
    \Delta i'(t) = \frac{ U_{\mathrm{L}} \Delta t}{L} = \frac{\hat u_{\mathrm{1}} \sin(\omega t) d(t) T_{\mathrm{s}}}{L} = \Delta i'_{\mathrm{1}}(t).
\end{equation}

Using $d(t) = 1 -\frac{1}{M} \sin(\omega t)$:
 \begin{equation}
     \Delta i'_{\mathrm{1}}(t) = \frac{U_{\mathrm{2}}\hat u_{\mathrm{1}}\sin (\omega t)}{U_{\mathrm{2}}L}(1-\frac{1}{M}\sin(\omega t)) T_{\mathrm{s}}\label{eq:ex05Deltai'}.
 \end{equation}
 Using $\frac{\hat u_{\mathrm{1}}}{U_{\mathrm{2}}} = \frac{1}{M}$, \eqref{eq:ex05Deltai'} becomes:
 \begin{equation}
     \Delta i'_{\mathrm{1}}(t) = \frac{U_{\mathrm{2}}T_{\mathrm{s}}}{LM}\sin (\omega t)(1-\frac{\sin(\omega t)}{M}). \label{eq:ex05Deltai1'}
 \end{equation}
   To be able to calculate the maximum value of the mains current
    $i_\mathrm{1}$, the following equation has to be set up from \eqref{eq:ex05Deltai1'} as:
    \begin{equation}
        \frac{U_{\mathrm{2}}T_{\mathrm{s}}}{L} \frac{\mathrm{d}}{\mathrm{d}t}\left[\frac{1}{M}\sin(\omega t)\left(1-\frac{1}{M}\right)\sin(\omega t)\right] =0.
    \end{equation}
    The first derivation is as follows:
    \begin{equation}
        \frac{-\omega \cos(\omega t)(2\sin(\omega t)-M)}{M^2}=0.\label{eq:ex05derivationi1'}
    \end{equation}
    If the term $\frac{M}{2}=\sin(\omega t)$ is applied to \eqref{eq:ex05derivationi1'} this becomes zero:

    \begin{equation}
        \frac{-\omega \cos(\omega t)(2 \frac{M}{2}-M)}{M^2}=0.
    \end{equation}
    The ratio $\frac{M}{2}=\sin(\omega t)$ can now be used in \eqref{eq:ex05Deltai1'} to calculate the maximum $\Delta i'_{\mathrm{1}}(t)$ as follows:
    \begin{equation}
        \Delta i'_{\mathrm{1}}(t) = \frac{U_{\mathrm{2}}T_{\mathrm{s}}M}{2LM}\left(1-\frac{M}{2M}\right) = \frac{U_{\mathrm{2}}\frac{1}{f_\mathrm{s}}}{4L}=\frac{\SI{400}{\volt}\cdot \SI{50}{\micro\s}}{4\cdot\SI{570}{\micro\henry}} = \SI{8.77}{\ampere}.
    \end{equation}
\end{solutionblock}

\subtask{Complete the current curve for a switching frequency $f_\mathrm{s2} = \SI{2}{\kilo\hertz}$ and a inductance $L = \SI{5}{\milli\henry}$ in \autoref{fig:Current i' control signal ex05}. Note: At the time $t =0$ is $i'=0$. The switch-on and switch-off times are determined by the control signal of the transistor $T$ and are summarized for the first 4 switching times in \autoref{tab:switching_times}.}

\begin{table}[ht]
    \centering
    
    \begin{tabular}{lcccc}
        \toprule
        & $i = 1$ & $2$ & $3$ & $4$ \\
        \midrule
       $T_\mathrm{i,OFF}$& \SI{480}{\micro\second} & \SI{935}{\micro\second} & \SI{1388}{\micro\second} & \SI{1852}{\micro\second} \\
        $T_\mathrm{i,ON}$  & \SI{500}{\micro\second} & \SI{1008}{\micro\second} & \SI{1520}{\micro\second} & \SI{2044}{\micro\second} \\
        \bottomrule
    \end{tabular}
    \caption{Switching times $T_\mathrm{i,OFF}$ and $T_\mathrm{i,ON}$ for different $i$-values.}
    \label{tab:switching_times}
\end{table}

\input{fig/ex05/Fig_VoltageCourses_UL_No2}
\begin{solutionblock}
    The component equation of inductance is used to determine the current values $i'(t)$ as:
    \begin{equation}
        \frac{\mathrm{d}}{\mathrm{d}t}i'(t) = \frac{1}{L}\hat u_\mathrm{1} \sin(\omega t).\label{eq:ex054Deltai'}
    \end{equation}
    \eqref{eq:ex054Deltai'} must now be converted to $i'$. To do this, the right-hand side of the equation must be integrated, with the initial state of the current $i'= 0$, to:
    \begin{equation}
        i'(t)=\frac{\hat u_\mathrm{1} \cos(\omega t)}{\omega L}.\label{eq:ex054currenti'1}
    \end{equation}
    The first interval is from $0 < t < T_\mathrm{1,off} = \SI{480}{\micro\s}$ for conducting transistor $T$. Inserting into \eqref{eq:ex054currenti'1} used with the upper and lower limit leads to:
    \begin{equation}
        i'(T_\mathrm{1,off}) = \frac{\hat u_\mathrm{1}}{L \omega}(1- \cos(\omega T_\mathrm{1,off})) = \frac{\sqrt{2}\cdot\SI{230}{\volt}}{\SI{5}{\milli\henry}\cdot 2\pi \cdot \SI{50}{\hertz}}\cdot (1-\cos(2\pi \cdot \SI{50}{\hertz} \cdot \SI{480}{\micro\s})) = \SI{2.35}{\ampere}.
    \end{equation}
    For the interval $T_\mathrm{1,off} < t < T_\mathrm{1,on} = \SI{500}{\micro\s}$ in which the diode $D_\mathrm{5}$ conducts, the voltage $U_\mathrm{L,1}$ is first determined as:
    \begin{equation}
        U_\mathrm{L,1} = u_\mathrm{1}(t) - U_\mathrm{2}= \hat u_\mathrm{1} \sin(\omega t) - U_\mathrm{2} = \sqrt{2} \cdot \SI{230}{\volt} \cdot \sin(2\pi \cdot \SI{50}{\hertz}\cdot \SI{20}{\micro\s}) - \SI{400}{\volt} = -\SI{350}{\volt}.
    \end{equation}
    This voltage value  $U_\mathrm{L,1}$ is used to calculate the current  $i'(T_\mathrm{1,on})$ as:
    \begin{equation}
        i'(T_\mathrm{1,on}) = i'(T_\mathrm{1,off}) -\frac{ U_\mathrm{L,1}}{L}(T_\mathrm{1,on}-T_\mathrm{1,off}) = \SI{2.35}{\ampere} -\frac{\SI{350}{\volt}}{\SI{5}{\milli\henry}}\cdot (\SI{500}{\micro\s}-\SI{480}{\micro\s}) = \SI{0.95}{\ampere}.
    \end{equation}
     The second interval  $T_\mathrm{1,on} < t < T_\mathrm{2,off}  = \SI{935}{\micro\s}$ in which the transistor $T$ conducts, have to be inserted  into \eqref{eq:ex054currenti'1} with the upper and lower limit. This leads to:
     \begin{equation}
        i'(T_\mathrm{2,off}) = \frac{\hat u_\mathrm{1}}{L \omega}(\cos (\omega T_\mathrm{1,on})- \cos(\omega T_\mathrm{2,off}))+ i'(T_\mathrm{1,on}).
    \end{equation}
    With values:
    \begin{equation}
        i'(T_\mathrm{2,off}) = \frac{\sqrt{2}\cdot\SI{230}{\volt}}{\SI{5}{\milli\henry}\cdot 2\pi \cdot \SI{50}{\hertz}}\cdot (\cos(2\pi \cdot \SI{50}{\hertz}\cdot \SI{500}{\micro\s})-\cos(2\pi \cdot \SI{50}{\hertz} \cdot \SI{935}{\micro\s})) + \SI{0.95}{\ampere}= \SI{7.27}{\ampere}.
     \end{equation}

     For the second interval $T_\mathrm{2,off} < t < T_\mathrm{2,on} = \SI{1008}{\micro\s}$ in which the diode $D_\mathrm{5}$ conducts, the voltage $U_\mathrm{L.2}$ is first determined as:
     \begin{equation}
        U_\mathrm{L,2} = u_\mathrm{1}(t) - U_\mathrm{2}= \hat u_\mathrm{1} \sin(\omega t) - U_\mathrm{2} = \sqrt{2} \cdot \SI{230}{\volt} \cdot \sin(2\pi \cdot \SI{50}{\hertz}\cdot \SI{73}{\micro\s}) - \SI{400}{\volt} = \SI{306}{\volt}.
    \end{equation}
    This voltage value  $U_\mathrm{L,2}$ is used to calculate the current  $i'(T_\mathrm{2,on})$ as:
    \begin{equation}
        i'(T_\mathrm{2,on}) = i'(T_\mathrm{2,off}) -\frac{ U_\mathrm{L,2}}{L}(T_\mathrm{2,on}-T_\mathrm{2,off}) = \SI{7.27}{\ampere} -\frac{\SI{306}{\volt}}{\SI{5}{\milli\henry}}\cdot (\SI{1008}{\micro\s}-\SI{935}{\micro\s}) = \SI{2.8}{\ampere}.
    \end{equation}
    The values determined in this task for mains current
    $i'(t)$ must be entered in \autoref{fig:Current i' control signal ex05}, just like the envelope. This results in the following diagram:
\end{solutionblock}

\input{fig/ex05/Fig_courses_i'_results}



\subtask{Approximately sketch the envelope of the current ripple in \autoref{fig:Current i' control signal ex05}. Sketch the voltage across the inductor in \autoref{fig:Voltage curve ul signal Transistor ex05} and enter the average value of the voltage as an approximation. How would the switch-on/switch-off ratio of the transistor have to be changed before and after the current peak in order to bring the average actual current value closer to the current setpoint?}
\begin{solutionblock}
    The envelope of the current ripple is scetched in \autoref{fig:Current i' control signal ex05 result}. Moreover the voltage across the inductor is scetched in \autoref{fig:Voltage curve ul signal Transistor ex05 res}. The average value of the voltage at the inductor is zero. Before the current peak in \autoref{fig:Current i' control signal ex05 result}, the transistor should be switched on for longer and then the switch-on phase should be reduced. This allows the current to be controlled better.
    \input{fig/ex05/Fig_VoltageCourses_UL_res.tex}
\end{solutionblock}
\subtask{Dimension the output capacitance $C$ in a way that the amplitude of the output voltage ripple is $\Delta u_{\mathrm{2}}<0.05  \hat u_{\mathrm{2}}$.
Note: assume an idealized lossless converter, with the assumption $u_2(t)\approx U_2$.}
\begin{solutionblock}
    For the input power $p_\mathrm{1}(t)$ follows $p_\mathrm{1}(t)= P_\mathrm{2} -\Delta p_\mathrm{2}$. The power at the load minus the power at the capacitor. The part $\Delta p_\mathrm{2}$ is zero on average, because it is the pulsating power across capacitor $C$. From this consideration it can be concluded for $p_\mathrm{1}(t)$:
    \begin{equation}
        p_\mathrm{1}(t) = \hat u_\mathrm{1} \sin(\omega t) \hat i_\mathrm{1} \sin(\omega t) = \hat u_\mathrm{1} \hat i_\mathrm{1} \sin^2(\omega t).
    \end{equation}
    By using the addition theorem $\sin^2(\omega t) = \frac{1}{2}(1-\cos(2\omega t))$ follows:
    \begin{equation}
        p_\mathrm{1}(t) = \frac{\hat u_\mathrm{1} \hat i_\mathrm{1}}{2}(1-\cos(2\omega t)).
    \end{equation}

    The power delivered across the load $R$ is equal to the output power $P_\mathrm{2}$ which is given as follows:
    \begin{equation}
        P_\mathrm{2} = \frac{\hat u_\mathrm{1}\hat i_\mathrm{1}}{2} = U_\mathrm{2} I_\mathrm{2}.
    \end{equation}
    The pulsating power across capacitor $C$ must be calculated as follows:
    \begin{equation}
        \Delta p_\mathrm{2} = \frac{\hat u_\mathrm{1}\hat i_\mathrm{1}}{2} \cos(2\omega t).
    \end{equation}
    The current through the capacitor is given by the following equation, using $P = UI$:
    \begin{equation}
        i_\mathrm{C}(t)=C \frac{\mathrm{d}u_\mathrm{c}}{\mathrm{d}t} = \frac{\Delta p_\mathrm{2}}{U_\mathrm{2}} = \frac{P_\mathrm{2}\cos(2\omega t)}{U_\mathrm{2}}.\label{eq:ex056currentic}
    \end{equation}
    The following follows from \eqref{eq:ex056currentic} for $U_\mathrm{C}$:
    \begin{equation}
        \frac{\mathrm{d}u_\mathrm{C}}{\mathrm{d}t} = \frac{P_\mathrm{2}\cos (2\omega t)}{C U_\mathrm{2}},
    \end{equation}

    \begin{equation}
        u_\mathrm{C} = \frac{1}{C} \int \frac{P_\mathrm{2}\cos(2\omega t)}{U_\mathrm{2}} \, \mathrm{d}t = \frac{P_\mathrm{2}\sin(2\omega t)}{2\omega C U_\mathrm{2}}.\label{eq:ex056voltageuc}
    \end{equation}
    $u_\mathrm{C}$ is equal to $ \hat u_\mathrm{C} \sin(2\omega t)$ from this follows for \eqref{eq:ex056voltageuc}:
    \begin{equation}
       \Delta u_\mathrm{2}= \hat u_\mathrm{C} = \frac{P_\mathrm{2}}{2\omega C U_\mathrm{2}}.
    \end{equation}
    And now the capacitance can be calculate with $\Delta u_{\mathrm{2}}<0.05  \hat u_{\mathrm{2}}$ as:
    \begin{equation}
        C = \frac{P_\mathrm{2}}{2\cdot 2\pi f  U_\mathrm{2}  \Delta u_\mathrm{2} }= \frac{U_\mathrm{2}I_\mathrm{2}}{2\cdot 2\pi f  U_\mathrm{2}\  0.05 U_\mathrm{2}} = \frac{\SI{400}{\volt}\cdot \SI{10}{\ampere}}{4 \pi \cdot\SI{50}{\hertz}\cdot\SI{400}{\volt}\cdot 0.05 \cdot \SI{400}{\volt}} = \SI{796}{\micro\farad}.  
    \end{equation}
    \end{solutionblock}
\subtask{Calculate the RMS value of the current through the capacitor $I_{\mathrm{C}}$.
Note: the mean value of the current through the capacitor is $\overline i_{\mathrm{C}}=0$!}
\begin{solutionblock}
    \begin{equation}
    i_\mathrm{2}(t) = i_\mathrm{C}(t)+i_\mathrm{R}(t).
\end{equation}
Using first biome:
\begin{equation}
    i^2_\mathrm{2}(t)= (i_\mathrm{C}(t)+i_\mathrm{R}(t))^2 = i^2_\mathrm{C}(t)+2i_\mathrm{R}(t) i_\mathrm{C}(t)+ i^2_\mathrm{R}(t).
\end{equation}
Calculate RMS value:
\begin{equation}
    I^2_\mathrm{2}(t)= \frac{1}{T_\mathrm{s}}\int_{0}^{T_\mathrm{s}}  i^2_\mathrm{2}(t) \, \mathrm{d}t\ = \frac{1}{T_\mathrm{s}}\int_{0}^{T_\mathrm{s}}  i^2_\mathrm{C}(t) \, \mathrm{d}t\ +\frac{1}{T_\mathrm{s}}\int_{0}^{T_\mathrm{s}}  2i_\mathrm{C}(t) i_\mathrm{R}(t)\, \mathrm{d}t\ +\frac{1}{T_\mathrm{s}}\int_{0}^{T_\mathrm{s}}  i^2_\mathrm{R}(t) \, \mathrm{d}t\
\end{equation}
Solving the equation, the term $\frac{1}{T_\mathrm{s}}\int_{0}^{T_\mathrm{s}} 2i_\mathrm{C}(t) i_\mathrm{R}(t)\, \mathrm{d}t$ is zero due to the periodicity of $i_\mathrm{C}(t)$:
\begin{equation}
    I^2_\mathrm{2}(t)=\frac{T_\mathrm{s}I^2_\mathrm{C}(t)}{T_\mathrm{s}}+\frac{T_\mathrm{s}I^2_\mathrm{R}(t)}{T_\mathrm{s}} =I^2_\mathrm{C}(t) +I^2_\mathrm{R}(t). \label{eq:ex056RMSIc}
\end{equation}
Finally, \eqref{eq:ex056RMSIc} must be converted to $I^2_\mathrm{C}(t)$:
\begin{equation}
    I^2_\mathrm{C}(t) = I^2_\mathrm{2}(t)-I^2_\mathrm{R}(t).%=\sqrt{\SI{209}{\ampere}-\SI{100}{\ampere}} = \SI{10.44}{\ampere}.
\end{equation}

\end{solutionblock}
\subtask{The current carrying capacity of the capacitor is $\frac{\SI{10}{\ampere}}{\SI{1}{\milli\farad}}$. How large must its capacitance be selected to stay below this threshold with $I_\mathrm{C}=\SI{10.44}{\ampere}$? Is the permissible output voltage fluctuation or is it the current carrying capacity that determines the capacitance?}
\begin{solutionblock}
    Calculation of the capacitance $C$:
    \begin{equation}
        C = \frac{I_\mathrm{C}}{\SI{10}{\frac{\ampere}{\milli\farad}}} = \frac{\SI{10.44}{\ampere}}{\SI{10}{\frac{\ampere}{\milli\farad}}}=\SI{1.04}{\milli\farad}.
    \end{equation}
    The dimensioned capacitance for an output voltage ripple $\Delta u_{\mathrm{2}}<0.05  \hat u_{\mathrm{2}}$ was determined as $\SI{796}{\micro\farad}$.
    If these two capacitor values are compared, it becomes clear that the current carrying capacity of the capacitor is the determining factor. 
\end{solutionblock}
\subtask{What is the current load (effective and average value) of the mains diodes?}
\begin{solutionblock}
    The rectification ensures that the current $\overline i_\mathrm{D,1-4}$ is calculated as the mean value over a half-period. So the average value is given with:
    \begin{equation}
        \overline{i}_{\mathrm{D,1-4}} = \frac{1}{2\pi} \int_{0}^{\pi} \hat{i}_{\mathrm{1}} \sin(\omega t) \, \mathrm{d}t\ = \frac{\hat{i}_{\mathrm{1}}}{2\pi} [-\cos(\pi)-[-\cos(0)]] = \frac{\hat{i}_{\mathrm{1}}}{\pi}.
    \end{equation}
    The effective value is given with:
    \begin{equation}
        I^2_{\mathrm{D,1-4}} = \frac{1}{2\pi} \int_{0}^{\pi} \hat{i}^2_{\mathrm{1}} \sin^2(\omega t) \, \mathrm{d}t\ = \frac{\hat{i}^2_{\mathrm{1}}}{2\pi}\left[\frac{\pi-\frac{\sin(2\pi)}{2}}{2}-\frac{\frac{-\sin(0)}{2}}{2}\right], 
    \end{equation}
    \begin{equation}
        I_{\mathrm{D,1-4}} = \sqrt{\frac{\hat{i}^2_{\mathrm{1}}}{2\pi}\frac{\pi}{2}} = \frac{\hat{i}^2_{\mathrm{1}}}{2}.
    \end{equation}
 \end{solutionblock}
% \input{fig/ex05/Fig_Current_Courses.tex}
